### 2019-03-21 tex/appendix/ellipse_intersection.tex
Initial Overleaf Import

In this appendix the intersection of two ellipses with the same shape parameters $(a,b) \in \R^2_{>0}$ is described with more detail, as well as determining the functions $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$ for two ellipses that intersect.

\section{Intersection}

Let $E_1$ and $E_2$ be two ellipses that the intersection will be determined here. Without loss of generality, let us assume that $E_1$ is at the origin and $E_2$ is located at the center $(h,k) \in \R^2$. Their equations are given by

\begin{align*}
    \frac{x^2}{a^2} + \frac{y^2}{b^2} = 1 && (E_1)\\
    \frac{(x-h)^2}{a^2} + \frac{(y-k)^2}{b^2} = 1 && (E_2)
\end{align*}

As they are both equal to one we can get the following

\begin{align*}
    b^2x^2 + a^2y^2 = b^2(x-h)^2 + a^2(y-k)^2 \\
    b^2(-2xh + h^2) + a^2(-2yk + k^2) = 0\\
    x(2hb^2) = b^2h^2 + a^2(-2yk + k^2)\\
    x = y\frac{-2yka^2}{2hb^2} + \frac{b^2h^2 + a^2k^2}{2hb^2}
\end{align*}

Which can be rewritten as

\begin{align*}
    x = y\alpha + \beta
\end{align*}

with the constants $\alpha$ and $\beta$ being

\begin{align*}
    \alpha = \frac{-2yka^2}{2hb^2} \\
    \beta = \frac{b^2h^2 + a^2k^2}{2hb^2}
\end{align*}

Then replacing it back to the equation of $E_1$ we get

\begin{align*}
    \frac{(y\alpha + \beta)^2}{a^2} + \frac{y^2}{b^2} = 1\\
    b^2(y\alpha + \beta)^2 + y^2a^2 - a^2b^2 = 0\\
    y^2(b^2\alpha^2 + a^2) + y(2\beta\alpha b^2) + b^2\beta^2 -a^2b^2 = 0
\end{align*}

Which is a second degree polynomial, therefore, $E_1$ and $E_2$ intersect if, and only if the roots of the polynomial are real. The intersection points itself can be obtained by solving the polynomial for $y$ and applying its value onto the $x=y\alpha + \beta$ equation.

\subsection{Determining $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$}

Let us assume that $E_1$ and $E_2$, each one with shape parameters $(a,b) \in \R^2_{>0}$, intersect at $p_1$ and $p_2$. Then, to determine $\Gamma_+(1,2)$ and $\Gamma_-(1,2)$, we need to first determine the angles of intersection of $p_1$ and $p_2$ on $E_1$. For that, let us first define another way that ellipses can be represented

\begin{definicao}
Let $E(q)$ be an ellipse with shape parameters $(a,b)\in \R^2_{>0}$, centered at $q \in \R^2$. Then, $E(q)$ is the set of points that are the image of curve $\gamma : [0, 2\pi] \mapsto \R^2$ given by

\begin{equation}\label{eq:parametric_ellipse}
\gamma(t) = \left\{
\begin{array}{l}
x(t)= a\cos{t} + q_x\\
y(t)=b\sin{t} + q_y
\end{array}
\right.
\end{equation}

\end{definicao}

From \autoref{eq:parametric_ellipse}, we can get that

\begin{align*}
\dfrac{y-q_y}{x-q_x} = \dfrac{b}{a}\tan{t}\\
t=\arctan\left(\dfrac{a}{b} \dfrac{y-q_y}{x-q_x}\right)
\end{align*}

However, as the image of $\arctan$ is $[-\frac{\pi}{2}, \frac{\pi}{2}]$, we need to check the sign of $x-q_x$ to determine the angle in $[0, 2\pi]$. After that, we can get the two angles that represent the intersection points on $E_1$.

To find out which one of the angles are $\Gamma_+(1,2)$, we need to go further and determine the derivative of $\gamma(t)$ which is going to be used to determine the vectors tangent to the ellipses at the intersection points.

\begin{equation}\label{eq:der_parametric_ellipse}
\gamma'(t) = \left\{
\begin{array}{l}
x(t)= -a\sin{t}\\
y(t)=b\cos{t}
\end{array}
\right.
\end{equation}

Let $\gamma_1$ and $\gamma_2$ be the curves describing $E_1$ and $E_2$ respectively. Also, let $s_1$ be the angle, such that $\gamma_1(s_1)=p_1$, and $t_1$ be the angle, such that $\gamma_2(t_1)=p_1$.
Then, the tangent vectors to the $E_1$ and $E_2$ at $p_1$ are $\gamma_1'(s_1)$ and $\gamma_2'(t_1)$ respectively.

\begin{figure}[H]
\centering

    \caption{Determining $\Gamma_+(1,2)$}
    \input{tex/figures/ellipse_gamma.tex}
    \fautor
    \label{fig:ellipse_gamma}
\end{figure}

The following lemma states a relation between $s_1$ and $\Gamma_+(1,2)$

\begin{lema}\label{lema:lema_ellipse_angle}
	Let $\theta$ be the angle between $\gamma_1'(s_1)$ and $\gamma_2'(t_1)$ (counter-clockwise). Then, $\theta > \pi$ if, and only if $\Gamma_+(1,2)=s_1$.
\end{lema}

Instead of a formal proof of \autoref{lema:lema_ellipse_angle}, a graphical explanation using \autoref{fig:ellipse_gamma} is provided.

First, let us state some facts that can also be seen in \autoref{fig:ellipse_gamma}

\begin{itemize}
    \item $E_1 \cap E_2$ is convex and bounded by two arcs, one from each ellipse.
    \item Starting at any of the intersection points, one of the $E_1 \cap E_2$ arcs will be clockwise-oriented and the other counter-clockwise-oriented.
    \item The counter-clockwise-oriented arc starting at $\Gamma_+(1,2)$ is from $E_1$.
\end{itemize}

Let us assume that $p_1$ is the intersection point which is the opening angle $\Gamma_+(1,2)$. Then, the vectors $\gamma_1'(s_1)$ and $-\gamma_2'(t_1)$ are tangent to the $E_1 \cap E_2$ area at point $p_1$. Because of the convexity of $E_1 \cap E_2$, the angle between $\gamma_1'(s_1)$ and $-\gamma_2'(t_1)$ has to be less than or equal to $\pi$, therefore the angle between $\gamma_1'(s_1)$ and $-\gamma_2'(t_1)$ has to be greater than $\pi$, which is what \autoref{lema:lema_ellipse_angle} says. It is easy to see that the converse is also true assuming that $p_1$ is the point which determines the angle $\Gamma_-(1,2)$.

Lastly, in \autoref{fig:ellipse_gamma}, it can be seen that if one point is classified as $\Gamma_+(1,2)$ the other will necessarily be classified as $\Gamma_-(1,2)$.
Computationally, this classification can be done taking the cross product of the two vectors and checking its signal, negative cross products imply an angle greater than $\pi$.