### 2019-03-21 tex/figures/ellipse_gamma.tex

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,193); %set diagram left start at 0, and has height of 193

%Shape: Ellipse [id:dp10634347053631776] 
\draw   (68.5,97.75) .. controls (68.5,67.24) and (115.62,42.5) .. (173.75,42.5) .. controls (231.88,42.5) and (279,67.24) .. (279,97.75) .. controls (279,128.26) and (231.88,153) .. (173.75,153) .. controls (115.62,153) and (68.5,128.26) .. (68.5,97.75) -- cycle ;
%Shape: Ellipse [id:dp7816950476685691] 
\draw  [color={rgb, 255:red, 163; green, 163; blue, 163 }  ,draw opacity=1 ] (167.5,103.75) .. controls (167.5,73.24) and (214.62,48.5) .. (272.75,48.5) .. controls (330.88,48.5) and (378,73.24) .. (378,103.75) .. controls (378,134.26) and (330.88,159) .. (272.75,159) .. controls (214.62,159) and (167.5,134.26) .. (167.5,103.75) -- cycle ;
%Straight Lines [id:da36505632658461096] 
\draw    (236,53) -- (185.87,33.72) ;
\draw [shift={(184,33)}, rotate = 381.03999999999996] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

%Straight Lines [id:da357569598162492] 
\draw    (234,52) -- (182.95,63.56) ;
\draw [shift={(181,64)}, rotate = 347.24] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

%Straight Lines [id:da9574900886541131] 
\draw    (212,149) -- (257.1,164.11) ;
\draw [shift={(259,164.75)}, rotate = 198.53] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

%Straight Lines [id:da5801314237314061] 
\draw    (212,149) -- (261.53,140.1) ;
\draw [shift={(263.5,139.75)}, rotate = 529.8199999999999] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]    (10.93,-3.29) .. controls (6.95,-1.4) and (3.31,-0.3) .. (0,0) .. controls (3.31,0.3) and (6.95,1.4) .. (10.93,3.29)   ;

%Shape: Arc [id:dp33725313902395126] 
\draw  [draw opacity=0][line width=0.75]  (219.73,152.25) .. controls (218.31,154.08) and (215.71,155.31) .. (212.73,155.31) .. controls (208.24,155.31) and (204.6,152.51) .. (204.6,149.06) .. controls (204.6,145.61) and (208.24,142.81) .. (212.73,142.81) .. controls (216.16,142.81) and (219.09,144.44) .. (220.29,146.74) -- (212.73,149.06) -- cycle ; \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]  (219.73,152.25) .. controls (218.31,154.08) and (215.71,155.31) .. (212.73,155.31) .. controls (208.24,155.31) and (204.6,152.51) .. (204.6,149.06) .. controls (204.6,145.61) and (208.24,142.81) .. (212.73,142.81) .. controls (216.16,142.81) and (219.09,144.44) .. (220.29,146.74) ;
%Shape: Arc [id:dp6533092564639917] 
\draw  [draw opacity=0] (216.22,56.17) .. controls (213.28,54.85) and (212,51.09) .. (213.38,47.64) .. controls (213.78,46.63) and (214.38,45.75) .. (215.09,45.04) -- (218.98,49.88) -- cycle ; \draw   (216.22,56.17) .. controls (213.28,54.85) and (212,51.09) .. (213.38,47.64) .. controls (213.78,46.63) and (214.38,45.75) .. (215.09,45.04) ;

% Text Node
\draw (216,26) node [scale=0.7]  {$\gamma _{1} '( s_{2})$};
% Text Node
\draw (164,70) node [scale=0.7]  {$\gamma _{2} '( t_{2})$};
% Text Node
\draw (202,50) node [scale=0.5]  {$\theta $};
% Text Node
\draw (284,143) node [scale=0.7]  {$\gamma _{1} '( s_{1})$};
% Text Node
\draw (286,170) node [scale=0.7]  {$\gamma _{2} '( t_{1})$};
% Text Node
\draw (206.67,136.67) node [scale=0.5]  {$\theta $};
% Text Node
\draw (214,163) node [scale=0.7]  {$p_{1}$};
% Text Node
\draw (237,41) node [scale=0.7]  {$p_{2}$};
% Text Node
\draw (56,93) node   {$E_{1}$};
% Text Node
\draw (390,97) node   {$E_{2}$};

\end{tikzpicture}