### 2019-03-22 tex/chapter/definicoes.tex

Some definitions and results that are used throughout the text are given in this chapter.

\section{Norm}

Norms are used in this chapter to define disks and ellipses.

\begin{definicao}
Let $\xi : \R^2 \mapsto [0, \infty]$, $\xi$ is said to be a norm function of $\R^2$ if for any $p, q \in \R^2$ and $a \in \R$,

\begin{enumerate}
    \item $\xi(p + q) \le \xi(p) + \xi(q)$
    \item $\xi(ap) = |a|\xi(p)$
    \item if $\xi(p)=0$, then $p=0$
\end{enumerate}

\end{definicao}

Probably the most known example of a norm function in $\R^2$ is the euclidean norm, described by \autoref{eq:norm2}.

\begin{equation}\label{eq:norm2}
||u||_2 = \sqrt{u^{T}u}
\end{equation}

A more general version of the euclidean norm is the here called elliptical norm which is defined as

\begin{definicao}
Let $Q$ be a $2$ by $2$ positive definite matrix, and $u \in \R^2$, then the elliptical norm is defined as

\begin{equation}
||u||_{elliptical} = \sqrt{u^{T}Qu}
\end{equation}
\end{definicao}

It is easy to see that the elliptical norm, when taking $Q$ to be the identity matrix, becomes the euclidean norm.

Given a norm function, determining the distance between two points is done by calculating the norm of the vector defined by the difference between the two points. For example, the elliptical distance between the points $p,q \in \R^2$ is $||p-q||_{elliptical}$.

\section{Disk}

A circle (or circumference) is a set of points in $\R^2$ that have constant euclidean distance, also known as radius, to another point, also referred to as the center of the circle. A unit circle is a circle with radius equal to $1$.
A disk is the set of points of a circle plus its interior, or in other words the set of points with euclidean distance to the center less than or equal the radius.

\section{Ellipse}

The ellipse is categorized, along with the parabola and the hyperbola, as a conic section. As the name suggests, conic sections are curves resulted from the intersection of a right circular cone in $\R^3$ with a plane \cite{brannan:geometry}. From that definition, an equation that describes any conic section is given as follows

\begin{equation}\label{equation:quadratic_form}
Ax^2 + Bxy + Cy^2 + Dx + Ey + F = 0
\end{equation}

Any ellipse, rotated or axis-parallel, can be described by \autoref{equation:quadratic_form}. To distinguish an ellipse from the other conic sections the condition $4AC - B^2>0$ can be verified \cite{ayoub}.

Then, assuming the center of an ellipse is $c \in \R^2$, then \autoref{equation:quadratic_form} can be rewritten as a quadratic form as follows

\begin{equation}
(p-c)^{T}Q(p-c) = 1
\end{equation}

with $p \in \R^2$ and $Q$ being a $2$ by $2$ positive definite matrix which carries the parameters of the ellipse. From \autoref{equation:quadratic_form}, $Q$ can be defined as follows

\[
Q=
\left( {\begin{array}{cc}
	A & \frac{B}{2} \\
	\frac{B}{2} & C \\
	\end{array} } \right).
\]

This makes us arrive at the following definition of the ellipse

\begin{definicao}\label{def:ellipse}
    Let $c\in \R^2$ be the center of an ellipse and $Q$ be a $2$ by $2$ positive definite matrix, an ellipse is the set of every point $p \in \R^2$ that have $||p-c||_{elliptical} = \sqrt{(p-c)^{T}Q(p-c)} = 1$. Also, a point $p$ is considered covered by an ellipse if $||p-c||_{elliptical} = \sqrt{(p-c)^{T}Q(p-c)} \le 1$.
\end{definicao}

An alternative way to define an ellipse, which can be seen as just a property derived from the definition above, is to begin the construction of the ellipse with two points called foci and a constant $R \in \R$, with $R$ being greater than the distance between the two foci points (see \autoref{fig:ellipse_with_foci}). The ellipse is, then, defined as the set of points whose distance to the foci is equal to $R$. In other words, let $f_1, f_2 \in \R^2$ be the two foci points, the ellipse is the set of every point $p \in \R^2$, such that $||p-f_1||_2 + ||p-f_2||_2 = R$. It can be shown that this definition is equivalent to \autoref{def:ellipse}.

\begin{figure}[H]
    \centering

    \caption{A non-axis-parallel ellipse and its foci points.}
    \input{tex/figures/ellipse_with_foci.tex}
    \fautor
    \label{fig:ellipse_with_foci}
\end{figure}

Also, in \autoref{fig:ellipse_with_foci}, the distance $a = ||p_a - c||_2$ is called the semi-major, and the distance $b = ||p_b-c||_2$ is called the semi-minor. These two values are also referred to as the shape parameters of an ellipse. Let $d = ||c-f_1||_2$, then it is easy to see that $a = R - d$ and $b = \sqrt{\frac{R^2}{4} - d^2}$.

Finally, an ellipse is said to be axis-parallel if its major-axis (see \autoref{fig:ellipse_with_foci}), which is the line that passes through its two foci points, is parallel to the $x$-axis.

\subsection{Axis parallel}

An axis parallel ellipse centered at $c = (c_x,c_y)$ can be described using \autoref{def:ellipse} with $Q$ being a diagonal matrix \footnote{the only non-zero terms are in the main diagonal}. This can be understood as a scaling transformation applied to the euclidean norm.
Defining the matrix $Q$ as

\[
Q=
\left( {\begin{array}{cc}
    \frac{1}{a^2} & 0 \\
    0 & \frac{1}{b^2} \\
    \end{array} } \right)
\]

Then, starting from \autoref{def:ellipse}, we can obtain the following equation

\begin{align*}
        (p-c)^{T}Q(p-c) = 1\\
    (\frac{p_x-c_x}{a^2}, \frac{p_y-c_y}{b^2})^{T}(p_x-c_x, p_y-c_y) = 1
 \end{align*}
 \begin{equation}
  \frac{(p_x-c_x)^2}{a^2} + \frac{(p_y-c_y)^2}{b^2} = 1
 \end{equation}

where $a$ and $b$ are the semi-major and semi-minor shape parameters respectively.

Another way to represent ellipses, which will be useful in some occasions, is through writing it as a curve, function of the angle with its major-axis (see \autoref{fig:ellipse_params}).

\begin{figure}[H]
    \centering

    \caption{The ellipse as a parametric curve}
    \input{tex/figures/ellipse_definition.tex}
    \fautor
    \label{fig:ellipse_params}
\end{figure}

Let $c \in \R^2$ be the center of an ellipse with shape parameters $(a,b) \in \R^2_{>0}$. Then $\gamma : [0, 2\pi] \mapsto \R^2$ defines a curve which maps every angle onto a point on the ellipse defined in \autoref{eq:parametric_ellipse}.

    \begin{equation}\label{eq:parametric_ellipse}
    \gamma(t) = \left\{
    \begin{array}{l}
    x(t)= a\cos{t} + c_x\\
    y(t)=b\sin{t} + c_y
    \end{array}
    \right.
    \end{equation}

No equivalent disk-circle wording exists for ellipses, this could be a source of ambiguity in the text, that is why a note for the reader was judged to be necessary. Throughout this work an ellipse will represent the set of points that satisfy \autoref{def:ellipse}. In some places, though, with prior clarification, we will denote as an ellipse, the set of points that are covered by the ellipse itself. For example, when we define $\Pp \cap E$ as the set of points in $\Pp$ that are covered by $E$, we are implicitly calling $E$ the set of points that are covered by the ellipse itself as it is defined by \autoref{def:ellipse}.