### 2019-04-29 quali/quali.presentation/figures/mcd_sim.tex

\usetikzlibrary{matrix}

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,95); %set diagram left start at 0, and has height of 95

%Shape: Circle [id:dp37777408295388804] 
\draw   (21,59.5) .. controls (21,45.69) and (32.19,34.5) .. (46,34.5) .. controls (59.81,34.5) and (71,45.69) .. (71,59.5) .. controls (71,73.31) and (59.81,84.5) .. (46,84.5) .. controls (32.19,84.5) and (21,73.31) .. (21,59.5) -- cycle ;
%Shape: Circle [id:dp4562198988843913] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (47.5,46.94) .. controls (47.5,33.14) and (58.69,21.94) .. (72.5,21.94) .. controls (86.31,21.94) and (97.5,33.14) .. (97.5,46.94) .. controls (97.5,60.75) and (86.31,71.94) .. (72.5,71.94) .. controls (58.69,71.94) and (47.5,60.75) .. (47.5,46.94) -- cycle ;
%Shape: Circle [id:dp8913971633985722] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (33,53.5) .. controls (33,39.69) and (44.19,28.5) .. (58,28.5) .. controls (71.81,28.5) and (83,39.69) .. (83,53.5) .. controls (83,67.31) and (71.81,78.5) .. (58,78.5) .. controls (44.19,78.5) and (33,67.31) .. (33,53.5) -- cycle ;
%Shape: Circle [id:dp05970043129919356] 
\draw  [color=green  ,draw opacity=1 ] (57.81,32.47) .. controls (57.81,18.66) and (69,7.47) .. (82.81,7.47) .. controls (96.62,7.47) and (107.81,18.66) .. (107.81,32.47) .. controls (107.81,46.27) and (96.62,57.47) .. (82.81,57.47) .. controls (69,57.47) and (57.81,46.27) .. (57.81,32.47) -- cycle ;
%Straight Lines [id:da04725899028195979] 
%\draw[densely dotted]  (21,59.5) -- (71,59.5) ;

%Flowchart: Extract [id:dp5267407418119328] 
\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (71,57.45) -- (72.52,59.55) -- (69.48,59.55) -- cycle ;

\draw (60,73) node [scale=0.5]  {$+$};
% Text Node
\draw (65.22,66) node [scale=0.5]  {$+$};
% Text Node
\draw (74.22,50) node [scale=0.5,color=red]  {$+$};
% Text Node
\draw (54.22,39.56) node [scale=0.5]  {$-$};
% Text Node
\draw (45.62,36.96) node [scale=0.5]  {$-$};
% Text Node
\draw (38.02,31.56) node [scale=0.5]  {$-$};

% Text Node
\draw (20,35) node [scale=0.5]  {$D_{1}$};
\end{tikzpicture}
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,95); %set diagram left start at 0, and has height of 95

%Shape: Circle [id:dp37777408295388804] 
\draw   (21,59.5) .. controls (21,45.69) and (32.19,34.5) .. (46,34.5) .. controls (59.81,34.5) and (71,45.69) .. (71,59.5) .. controls (71,73.31) and (59.81,84.5) .. (46,84.5) .. controls (32.19,84.5) and (21,73.31) .. (21,59.5) -- cycle ;
%Shape: Circle [id:dp4562198988843913] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (47.5,46.94) .. controls (47.5,33.14) and (58.69,21.94) .. (72.5,21.94) .. controls (86.31,21.94) and (97.5,33.14) .. (97.5,46.94) .. controls (97.5,60.75) and (86.31,71.94) .. (72.5,71.94) .. controls (58.69,71.94) and (47.5,60.75) .. (47.5,46.94) -- cycle ;
%Shape: Circle [id:dp8913971633985722] 
\draw  [color=green  ,draw opacity=1 ] (33,53.5) .. controls (33,39.69) and (44.19,28.5) .. (58,28.5) .. controls (71.81,28.5) and (83,39.69) .. (83,53.5) .. controls (83,67.31) and (71.81,78.5) .. (58,78.5) .. controls (44.19,78.5) and (33,67.31) .. (33,53.5) -- cycle ;
%Shape: Circle [id:dp05970043129919356] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (57.81,32.47) .. controls (57.81,18.66) and (69,7.47) .. (82.81,7.47) .. controls (96.62,7.47) and (107.81,18.66) .. (107.81,32.47) .. controls (107.81,46.27) and (96.62,57.47) .. (82.81,57.47) .. controls (69,57.47) and (57.81,46.27) .. (57.81,32.47) -- cycle ;
%Straight Lines [id:da04725899028195979] 
%\draw[densely dotted]  (21,59.5) -- (71,59.5) ;

%Flowchart: Extract [id:dp5267407418119328] 
\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (71,57.45) -- (72.52,59.55) -- (69.48,59.55) -- cycle ;

\draw (60,73) node [scale=0.5,color=red]  {$+$};
% Text Node
\draw (65.22,66) node [scale=0.5]  {$+$};
% Text Node
\draw (74.22,50) node [scale=0.5]  {$+$};
% Text Node
\draw (54.22,39.56) node [scale=0.5]  {$-$};
% Text Node
\draw (45.62,36.96) node [scale=0.5]  {$-$};
% Text Node
\draw (38.02,31.56) node [scale=0.5]  {$-$};

% Text Node
\draw (20,35) node [scale=0.5]  {$D_{1}$};
\end{tikzpicture}
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,95); %set diagram left start at 0, and has height of 95

%Shape: Circle [id:dp37777408295388804] 
\draw   (21,59.5) .. controls (21,45.69) and (32.19,34.5) .. (46,34.5) .. controls (59.81,34.5) and (71,45.69) .. (71,59.5) .. controls (71,73.31) and (59.81,84.5) .. (46,84.5) .. controls (32.19,84.5) and (21,73.31) .. (21,59.5) -- cycle ;

%Shape: Circle [id:dp4562198988843913] 
\draw  [color=green  ,draw opacity=1 ] (47.5,46.94) .. controls (47.5,33.14) and (58.69,21.94) .. (72.5,21.94) .. controls (86.31,21.94) and (97.5,33.14) .. (97.5,46.94) .. controls (97.5,60.75) and (86.31,71.94) .. (72.5,71.94) .. controls (58.69,71.94) and (47.5,60.75) .. (47.5,46.94) -- cycle ;

%Shape: Circle [id:dp8913971633985722] 
\draw  [color=green  ,draw opacity=1 ] (33,53.5) .. controls (33,39.69) and (44.19,28.5) .. (58,28.5) .. controls (71.81,28.5) and (83,39.69) .. (83,53.5) .. controls (83,67.31) and (71.81,78.5) .. (58,78.5) .. controls (44.19,78.5) and (33,67.31) .. (33,53.5) -- cycle ;

%Shape: Circle [id:dp05970043129919356] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (57.81,32.47) .. controls (57.81,18.66) and (69,7.47) .. (82.81,7.47) .. controls (96.62,7.47) and (107.81,18.66) .. (107.81,32.47) .. controls (107.81,46.27) and (96.62,57.47) .. (82.81,57.47) .. controls (69,57.47) and (57.81,46.27) .. (57.81,32.47) -- cycle ;
%Straight Lines [id:da04725899028195979] 
%\draw[densely dotted]  (21,59.5) -- (71,59.5) ;

%Flowchart: Extract [id:dp5267407418119328] 
\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (71,57.45) -- (72.52,59.55) -- (69.48,59.55) -- cycle ;

\draw (60,73) node [scale=0.5]  {$+$};
% Text Node
\draw (65.22,66) node [scale=0.5,color=red]  {$+$};
% Text Node
\draw (74.22,50) node [scale=0.5]  {$+$};
% Text Node
\draw (54.22,39.56) node [scale=0.5]  {$-$};
% Text Node
\draw (45.62,36.96) node [scale=0.5]  {$-$};
% Text Node
\draw (38.02,31.56) node [scale=0.5]  {$-$};

% Text Node
\draw (20,35) node [scale=0.5]  {$D_{1}$};
\end{tikzpicture}
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,95); %set diagram left start at 0, and has height of 95

%Shape: Circle [id:dp37777408295388804] 
\draw   (21,59.5) .. controls (21,45.69) and (32.19,34.5) .. (46,34.5) .. controls (59.81,34.5) and (71,45.69) .. (71,59.5) .. controls (71,73.31) and (59.81,84.5) .. (46,84.5) .. controls (32.19,84.5) and (21,73.31) .. (21,59.5) -- cycle ;

%Shape: Circle [id:dp4562198988843913] 
\draw  [color=green  ,draw opacity=1 ] (47.5,46.94) .. controls (47.5,33.14) and (58.69,21.94) .. (72.5,21.94) .. controls (86.31,21.94) and (97.5,33.14) .. (97.5,46.94) .. controls (97.5,60.75) and (86.31,71.94) .. (72.5,71.94) .. controls (58.69,71.94) and (47.5,60.75) .. (47.5,46.94) -- cycle ;

%Shape: Circle [id:dp8913971633985722] 
\draw  [color=green  ,draw opacity=1 ] (33,53.5) .. controls (33,39.69) and (44.19,28.5) .. (58,28.5) .. controls (71.81,28.5) and (83,39.69) .. (83,53.5) .. controls (83,67.31) and (71.81,78.5) .. (58,78.5) .. controls (44.19,78.5) and (33,67.31) .. (33,53.5) -- cycle ;

%Shape: Circle [id:dp05970043129919356] 
\draw  [color=green ,draw opacity=1 ] (57.81,32.47) .. controls (57.81,18.66) and (69,7.47) .. (82.81,7.47) .. controls (96.62,7.47) and (107.81,18.66) .. (107.81,32.47) .. controls (107.81,46.27) and (96.62,57.47) .. (82.81,57.47) .. controls (69,57.47) and (57.81,46.27) .. (57.81,32.47) -- cycle ;
%Straight Lines [id:da04725899028195979] 
%\draw[densely dotted]  (21,59.5) -- (71,59.5) ;

%Flowchart: Extract [id:dp5267407418119328] 
\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (71,57.45) -- (72.52,59.55) -- (69.48,59.55) -- cycle ;

\draw (60,73) node [scale=0.5]  {$+$};
% Text Node
\draw (65.22,66) node [scale=0.5]  {$+$};
% Text Node
\draw (74.22,50) node [scale=0.5,color=red]  {$+$};
% Text Node
\draw (54.22,39.56) node [scale=0.5]  {$-$};
% Text Node
\draw (45.62,36.96) node [scale=0.5]  {$-$};
% Text Node
\draw (38.02,31.56) node [scale=0.5]  {$-$};

% Text Node
\draw (20,35) node [scale=0.5]  {$D_{1}$};
\end{tikzpicture}
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,95); %set diagram left start at 0, and has height of 95

%Shape: Circle [id:dp37777408295388804] 
\draw   (21,59.5) .. controls (21,45.69) and (32.19,34.5) .. (46,34.5) .. controls (59.81,34.5) and (71,45.69) .. (71,59.5) .. controls (71,73.31) and (59.81,84.5) .. (46,84.5) .. controls (32.19,84.5) and (21,73.31) .. (21,59.5) -- cycle ;

%Shape: Circle [id:dp4562198988843913] 
\draw  [color=green  ,draw opacity=1 ] (47.5,46.94) .. controls (47.5,33.14) and (58.69,21.94) .. (72.5,21.94) .. controls (86.31,21.94) and (97.5,33.14) .. (97.5,46.94) .. controls (97.5,60.75) and (86.31,71.94) .. (72.5,71.94) .. controls (58.69,71.94) and (47.5,60.75) .. (47.5,46.94) -- cycle ;

%Shape: Circle [id:dp8913971633985722] 
\draw  [color=green  ,draw opacity=1 ] (33,53.5) .. controls (33,39.69) and (44.19,28.5) .. (58,28.5) .. controls (71.81,28.5) and (83,39.69) .. (83,53.5) .. controls (83,67.31) and (71.81,78.5) .. (58,78.5) .. controls (44.19,78.5) and (33,67.31) .. (33,53.5) -- cycle ;

%Shape: Circle [id:dp05970043129919356] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (57.81,32.47) .. controls (57.81,18.66) and (69,7.47) .. (82.81,7.47) .. controls (96.62,7.47) and (107.81,18.66) .. (107.81,32.47) .. controls (107.81,46.27) and (96.62,57.47) .. (82.81,57.47) .. controls (69,57.47) and (57.81,46.27) .. (57.81,32.47) -- cycle ;
%Straight Lines [id:da04725899028195979] 
%\draw[densely dotted]  (21,59.5) -- (71,59.5) ;

%Flowchart: Extract [id:dp5267407418119328] 
\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (71,57.45) -- (72.52,59.55) -- (69.48,59.55) -- cycle ;

\draw (60,73) node [scale=0.5]  {$+$};
% Text Node
\draw (65.22,66) node [scale=0.5]  {$+$};
% Text Node
\draw (74.22,50) node [scale=0.5]  {$+$};
% Text Node
\draw (54.22,39.56) node [scale=0.5,color=red]  {$-$};
% Text Node
\draw (45.62,36.96) node [scale=0.5]  {$-$};
% Text Node
\draw (38.02,31.56) node [scale=0.5]  {$-$};

% Text Node
\draw (20,35) node [scale=0.5]  {$D_{1}$};
\end{tikzpicture}
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,95); %set diagram left start at 0, and has height of 95

%Shape: Circle [id:dp37777408295388804] 
\draw   (21,59.5) .. controls (21,45.69) and (32.19,34.5) .. (46,34.5) .. controls (59.81,34.5) and (71,45.69) .. (71,59.5) .. controls (71,73.31) and (59.81,84.5) .. (46,84.5) .. controls (32.19,84.5) and (21,73.31) .. (21,59.5) -- cycle ;

%Shape: Circle [id:dp4562198988843913] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 },draw opacity=1 ] (47.5,46.94) .. controls (47.5,33.14) and (58.69,21.94) .. (72.5,21.94) .. controls (86.31,21.94) and (97.5,33.14) .. (97.5,46.94) .. controls (97.5,60.75) and (86.31,71.94) .. (72.5,71.94) .. controls (58.69,71.94) and (47.5,60.75) .. (47.5,46.94) -- cycle ;

%Shape: Circle [id:dp8913971633985722] 
\draw  [color=green  ,draw opacity=1 ] (33,53.5) .. controls (33,39.69) and (44.19,28.5) .. (58,28.5) .. controls (71.81,28.5) and (83,39.69) .. (83,53.5) .. controls (83,67.31) and (71.81,78.5) .. (58,78.5) .. controls (44.19,78.5) and (33,67.31) .. (33,53.5) -- cycle ;

%Shape: Circle [id:dp05970043129919356] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (57.81,32.47) .. controls (57.81,18.66) and (69,7.47) .. (82.81,7.47) .. controls (96.62,7.47) and (107.81,18.66) .. (107.81,32.47) .. controls (107.81,46.27) and (96.62,57.47) .. (82.81,57.47) .. controls (69,57.47) and (57.81,46.27) .. (57.81,32.47) -- cycle ;
%Straight Lines [id:da04725899028195979] 
%\draw[densely dotted]  (21,59.5) -- (71,59.5) ;

%Flowchart: Extract [id:dp5267407418119328] 
\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (71,57.45) -- (72.52,59.55) -- (69.48,59.55) -- cycle ;

\draw (60,73) node [scale=0.5]  {$+$};
% Text Node
\draw (65.22,66) node [scale=0.5]  {$+$};
% Text Node
\draw (74.22,50) node [scale=0.5]  {$+$};
% Text Node
\draw (54.22,39.56) node [scale=0.5]  {$-$};
% Text Node
\draw (45.62,36.96) node [scale=0.5,color=red]  {$-$};
% Text Node
\draw (38.02,31.56) node [scale=0.5]  {$-$};

% Text Node
\draw (20,35) node [scale=0.5]  {$D_{1}$};
\end{tikzpicture}
\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,95); %set diagram left start at 0, and has height of 95

%Shape: Circle [id:dp37777408295388804] 
\draw   (21,59.5) .. controls (21,45.69) and (32.19,34.5) .. (46,34.5) .. controls (59.81,34.5) and (71,45.69) .. (71,59.5) .. controls (71,73.31) and (59.81,84.5) .. (46,84.5) .. controls (32.19,84.5) and (21,73.31) .. (21,59.5) -- cycle ;

%Shape: Circle [id:dp4562198988843913] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 },draw opacity=1 ] (47.5,46.94) .. controls (47.5,33.14) and (58.69,21.94) .. (72.5,21.94) .. controls (86.31,21.94) and (97.5,33.14) .. (97.5,46.94) .. controls (97.5,60.75) and (86.31,71.94) .. (72.5,71.94) .. controls (58.69,71.94) and (47.5,60.75) .. (47.5,46.94) -- cycle ;

%Shape: Circle [id:dp8913971633985722] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 } ,draw opacity=1 ] (33,53.5) .. controls (33,39.69) and (44.19,28.5) .. (58,28.5) .. controls (71.81,28.5) and (83,39.69) .. (83,53.5) .. controls (83,67.31) and (71.81,78.5) .. (58,78.5) .. controls (44.19,78.5) and (33,67.31) .. (33,53.5) -- cycle ;

%Shape: Circle [id:dp05970043129919356] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] (57.81,32.47) .. controls (57.81,18.66) and (69,7.47) .. (82.81,7.47) .. controls (96.62,7.47) and (107.81,18.66) .. (107.81,32.47) .. controls (107.81,46.27) and (96.62,57.47) .. (82.81,57.47) .. controls (69,57.47) and (57.81,46.27) .. (57.81,32.47) -- cycle ;
%Straight Lines [id:da04725899028195979] 
%\draw[densely dotted]  (21,59.5) -- (71,59.5) ;

%Flowchart: Extract [id:dp5267407418119328] 
\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (71,57.45) -- (72.52,59.55) -- (69.48,59.55) -- cycle ;

\draw (60,73) node [scale=0.5]  {$+$};
% Text Node
\draw (65.22,66) node [scale=0.5]  {$+$};
% Text Node
\draw (74.22,50) node [scale=0.5]  {$+$};
% Text Node
\draw (54.22,39.56) node [scale=0.5]  {$-$};
% Text Node
\draw (45.62,36.96) node [scale=0.5]  {$-$};
% Text Node
\draw (38.02,31.56) node [scale=0.5,color=red]  {\textbf{$-$}};

% Text Node
\draw (20,35) node [scale=0.5]  {$D_{1}$};
\end{tikzpicture}