### 2019-05-04 quali/quali.presentation/figures/example_2_ellipses.tex
wip

\begin{tikzpicture}[scale=0.8]
\begin{axis}[height = {101.6mm}, axis equal = {true}, ylabel = {}, xmin = {-0.04304650080536887}, xmax = {5.015893955055976}, ymax = {4.953655933510717}, xlabel = {}, unbounded coords=jump,scaled x ticks = false,xlabel style = {font = {\fontsize{11 pt}{14.3 pt}\selectfont}, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0, rotate = 0.0},xmajorgrids = true,xtick = {0.0,1.0,2.0,3.0,4.0,5.0},xticklabels = {},xtick align = inside,xticklabel style = {font = {\fontsize{8 pt}{10.4 pt}\selectfont}, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0, rotate = 0.0},x grid style = {color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 0.1,
line width = 0.5,
solid},axis lines* = left,separate axis lines,x axis line style = {draw opacity = 0},scaled y ticks = false,ylabel style = {font = {\fontsize{11 pt}{14.3 pt}\selectfont}, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0, rotate = 0.0},ymajorgrids = true,ytick = {0.0,1.0,2.0,3.0,4.0},yticklabels = {},ytick align = inside,yticklabel style = {font = {\fontsize{8 pt}{10.4 pt}\selectfont}, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0, rotate = 0.0},y grid style = {color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 0.1,
line width = 0.5,
solid},axis lines* = left,separate axis lines,y axis line style = {draw opacity = 0},    xshift = 0.0mm,
    yshift = 0.0mm,
    axis background/.style={fill={rgb,1:red,1.00000000;green,1.00000000;blue,1.00000000}}
,legend style = {color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 1,
solid,fill = {rgb,1:red,1.00000000;green,1.00000000;blue,1.00000000},font = {\fontsize{8 pt}{10.4 pt}\selectfont}},colorbar style={title=}, ymin = {-0.025453315895344608}, width = {152.4mm}]\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.602918949315934, 0.1421191370308239)
};
\addplot+[draw=none, color = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.396020400537423, 2.2686705873010817)
};
\addplot+[draw=none, color = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(4.872716394984428, 3.773817627065047)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(3.202948490978642, 1.6235525016076402)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.848573275272137, 0.20375192853344104)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(1.8690841880778053, 1.6568512573232197)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.300100178756294, 4.05463630589837)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(1.6738171002119706, 3.3074971831952427)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.907448354814198, 4.598195661337079)
};
\addplot+[draw=none, color = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(0.1001310592661786, 3.3500841191834887)
};
\addplot+[draw=none, color = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(0.9130732329791647, 3.4401427131346485)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(3.4385814568108195, 4.065543992107185)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.354854569358288, 0.5753620833645212)
};
\addplot+[draw=none, color = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(4.393915159418624, 4.812737747206772)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(3.50505217637773, 0.20556942673409795)
};
\addplot+[draw=none, color = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,1.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(3.427160885660112, 2.538013336230315)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.1063177109467004, 4.409416548982065)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(2.3672535618848976, 1.8968192015410446)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(3.213442183570654, 2.9502323475574324)
};
\addplot+[draw=none, color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000},
draw opacity = 1.0,
line width = 0,
solid,mark = *,
mark size = 1.25,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, fill opacity = 1.0,
    line width = 0.5,
    rotate = 0,
    solid
},forget plot] coordinates {
(3.1034740796888682, 3.8975002907243397)
};
\addplot+ [color = {rgb,1:red,0.88887350;green,0.43564919;blue,0.27812294},
draw opacity = 1.0,
line width = 1,
solid,mark = none,
mark size = 2.0,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.88887350;green,0.43564919;blue,0.27812294}, fill opacity = 1.0,
    line width = 1,
    rotate = 0,
    solid
}]coordinates {
(3.5821391801246905, 3.725768625135797)
(3.581640769250864, 3.7573371749006075)
(3.5801460334561837, 3.7888742564484703)
(3.5776564627257965, 3.8203484329306416)
(3.574174538717948, 3.8517283302035143)
(3.5697037322902143, 3.8829826681030477)
(3.564248500039671, 3.9140802916255146)
(3.557814279860464, 3.9449902019834884)
(3.5504074855231975, 3.9756815875061053)
(3.5420355002815476, 4.006123854352811)
(3.532706669512473, 4.036286657009966)
(3.522430292397366, 4.06613992853991)
(3.511216612652421, 4.095653910552343)
(3.4990768083174792, 4.124799182868138)
(3.486022980613514, 4.153546692846007)
(3.4720681418798707, 4.1818677843428125)
(3.4572262026032843, 4.209734226278636)
(3.4415119575516027, 4.237118240778123)
(3.424941071026041, 4.263992530860086)
(3.407530061246667, 4.290330307647715)
(3.389296283886676, 4.316105317072325)
(3.3702579147718827, 4.341291866043976)
(3.3504339317626615, 4.365864848062904)
(3.3298440958363997, 4.389799768246228)
(3.30850893138933, 4.413072767744981)
(3.2864497057773625, 4.43566064752713)
(3.2636884081163244, 4.457540891502873)
(3.2402477273627284, 4.478691688969174)
(3.2161510296969293, 4.499091956351136)
(3.191422335231207, 4.518721358218576)
(3.166086294065997, 4.537560327556817)
(3.1401681617181305, 4.555590085271523)
(3.1136937739455917, 4.572792658908096)
(3.086689520993868, 4.58915090056702)
(3.05918232128958, 4.604648503997257)
(3.0311995946076107, 4.619270020850671)
(3.0027692347384747, 4.633000876081278)
(2.973919581683184, 4.645827382473971)
(2.944679393403315, 4.657736754288232)
(2.9150778171544514, 4.668717120003234)
(2.885144360431563, 4.678757534151636)
(2.8549088615552938, 4.68784798823026)
(2.8244014599284735, 4.6959794206767835)
(2.7936525659925096, 4.703143725902504)
(2.762692830913593, 4.709333762372166)
(2.7315531160289512, 4.7145433597228)
(2.7002644620835987, 4.718767324914467)
(2.6688580582882415, 4.722001447406804)
(2.6373652112291985, 4.724242503356176)
(2.6058173136613147, 4.725488258829275)
(2.574245813214977, 4.725737472029953)
(2.5426821810484377, 4.724989894537073)
(2.5111578804766745, 4.723246271552136)
(2.4797043356080772, 4.720508341156454)
(2.4483529000202116, 4.716778832578589)
(2.4171348255058915, 4.7120614634738)
(2.3860812309207127, 4.706360936218201)
(2.3552230711631004, 4.699682933221335)
(2.3245911063177935, 4.692034111261819)
(2.294215870993524, 4.683422094851727)
(2.2641276438854523, 4.673855468636306)
(2.234356417592708, 4.663343768836622)
(2.2049318687211144, 4.651897473743638)
(2.175883328300902, 4.639527993273234)
(2.1472397525488973, 4.626247657592549)
(2.1190296940043427, 4.612069704829006)
(2.091281273067097, 4.597008267874257)
(2.0640221499666134, 4.581078360296209)
(2.03727949718962, 4.564295861373174)
(2.0110799723939956, 4.5466775002650595)
(1.9854496918358344, 4.528240839337375)
(1.9604142043361954, 4.509004256654688)
(1.9359984658134808, 4.4889869276609655)
(1.9122268144068353, 4.46820880606508)
(1.8891229462153587, 4.446690603950513)
(1.8667098916773193, 4.424453771129103)
(1.8450099926129115, 4.401520473759406)
(1.824044879953445, 4.377913572250984)
(1.8038354521791604, 4.353656598476655)
(1.7844018544871711, 4.328773732315412)
(1.7657634587102913, 4.303289777549385)
(1.747938844006773, 4.277230137138904)
(1.730945778340196, 4.2506207879002655)
(1.714801200767976, 4.22348825461148)
(1.6995212045561434, 4.1958595835718)
(1.685121021137227, 4.167762315641376)
(1.6716150049272283, 4.139224458787931)
(1.6590166190168298, 4.110274460167808)
(1.6473384217510922, 4.080941177769225)
(1.6365920542110235, 4.051253851646009)
(1.6267882286094957, 4.021242074770467)
(1.6179367176130788, 3.9909357635344755)
(1.6100463446004336, 3.9603651279281658)
(1.603124974866976, 3.929560641425949)
(1.59717950778458, 3.8985530106098967)
(1.5922158699241333, 3.8673731445607484)
(1.5882390091478027, 3.8360521240470717)
(1.5852528896768971, 3.8046211705432733)
(1.583260488140247, 3.773111615107355)
(1.5822637906070331, 3.7415548671494343)
(1.5822637906070331, 3.70998238312216)
(1.583260488140247, 3.678425635164239)
(1.5852528896768971, 3.6469160797283204)
(1.5882390091478027, 3.615485126224522)
(1.592215869924133, 3.5841641057108458)
(1.5971795077845798, 3.552984239661697)
(1.6031249748669758, 3.5219766088456446)
(1.6100463446004336, 3.4911721223434284)
(1.6179367176130786, 3.460601486737118)
(1.6267882286094957, 3.430295175501127)
(1.6365920542110233, 3.4002833986255854)
(1.6473384217510922, 3.3705960725023685)
(1.6590166190168296, 3.3412627901037864)
(1.6716150049272283, 3.312312791483663)
(1.6851210211372267, 3.283774934630218)
(1.6995212045561434, 3.255677666699794)
(1.7148012007679758, 3.2280489956601137)
(1.7309457783401958, 3.2009164623713287)
(1.747938844006773, 3.1743071131326897)
(1.7657634587102913, 3.1482474727222085)
(1.784401854487171, 3.122763517956183)
(1.8038354521791602, 3.097880651794939)
(1.824044879953445, 3.0736236780206103)
(1.8450099926129115, 3.050016776512188)
(1.866709891677319, 3.0270834791424903)
(1.8891229462153585, 3.004846646321081)
(1.912226814406835, 2.983328444206514)
(1.9359984658134803, 2.9625503226106287)
(1.9604142043361952, 2.9425329936169065)
(1.985449691835834, 2.923296410934219)
(2.0110799723939956, 2.9048597500065343)
(2.03727949718962, 2.8872413888984196)
(2.064022149966613, 2.870458889975385)
(2.091281273067097, 2.8545289823973374)
(2.1190296940043423, 2.8394675454425884)
(2.1472397525488973, 2.825289592679045)
(2.175883328300902, 2.81200925699836)
(2.204931868721114, 2.7996397765279557)
(2.234356417592708, 2.7881934814349716)
(2.264127643885452, 2.7776817816352875)
(2.294215870993524, 2.7681151554198675)
(2.324591106317794, 2.759503139009775)
(2.3552230711631, 2.751854317050259)
(2.3860812309207122, 2.745176314053393)
(2.4171348255058906, 2.739475786797794)
(2.448352900020211, 2.734758417693005)
(2.4797043356080772, 2.73102890911514)
(2.511157880476674, 2.7282909787194582)
(2.5426821810484377, 2.726547355734521)
(2.5742458132149766, 2.7257997782416403)
(2.6058173136613143, 2.726048991442319)
(2.637365211229199, 2.727294746915418)
(2.668858058288241, 2.72953580286479)
(2.7002644620835987, 2.732769925357127)
(2.731553116028951, 2.7369938905487943)
(2.7626928309135925, 2.742203487899427)
(2.7936525659925096, 2.7483935243690896)
(2.8244014599284735, 2.7555578295948107)
(2.8549088615552938, 2.763689262041334)
(2.8851443604315627, 2.7727797161199574)
(2.9150778171544514, 2.7828201302683597)
(2.944679393403314, 2.793800495983362)
(2.9739195816831834, 2.8057098677976224)
(3.0027692347384747, 2.8185363741903156)
(3.0311995946076102, 2.8322672294209226)
(3.05918232128958, 2.8468887462743364)
(3.0866895209938674, 2.862386349704573)
(3.1136937739455917, 2.878744591363498)
(3.1401681617181305, 2.895947165000071)
(3.1660862940659964, 2.913976922714776)
(3.191422335231207, 2.9328158920530183)
(3.2161510296969285, 2.9524452939204573)
(3.240247727362728, 2.9728455613024196)
(3.2636884081163244, 2.99399635876872)
(3.2864497057773625, 3.0158766027444637)
(3.30850893138933, 3.0384644825266127)
(3.3298440958363993, 3.0617374820253653)
(3.350433931762661, 3.0856724022086897)
(3.3702579147718827, 3.1102453842276176)
(3.389296283886676, 3.135431933199268)
(3.407530061246667, 3.161206942623879)
(3.4249410710260406, 3.187544719411508)
(3.441511957551602, 3.21441900949347)
(3.4572262026032843, 3.2418030239929583)
(3.4720681418798707, 3.2696694659287804)
(3.486022980613514, 3.2979905574255874)
(3.4990768083174792, 3.3267380674034555)
(3.511216612652421, 3.35588333971925)
(3.522430292397366, 3.3853973217316833)
(3.532706669512473, 3.4152505932616277)
(3.5420355002815476, 3.4454133959187825)
(3.5504074855231975, 3.475855662765488)
(3.557814279860464, 3.5065470482881054)
(3.5642485000396706, 3.5374569586460782)
(3.569703732290214, 3.568554582168546)
(3.574174538717948, 3.5998089200680794)
(3.5776564627257965, 3.6311888173409517)
(3.5801460334561837, 3.662662993823123)
(3.581640769250864, 3.694200075370986)
(3.5821391801246905, 3.7257686251357964)
};
\addlegendentry{$E_1$}
\addplot+ [color = {rgb,1:red,0.24222430;green,0.64327509;blue,0.30444865},
draw opacity = 1.0,
line width = 1,
solid,mark = none,
mark size = 2.0,
mark options = {
    color = {rgb,1:red,0.00000000;green,0.00000000;blue,0.00000000}, draw opacity = 1.0,
    fill = {rgb,1:red,0.24222430;green,0.64327509;blue,0.30444865}, fill opacity = 1.0,
    line width = 1,
    rotate = 0,
    solid
}]coordinates {
(4.324183803520279, 1.1154337173027569)
(4.323486028296922, 1.1470022670675675)
(4.3213933981843695, 1.1785393486154305)
(4.317907999161827, 1.2100135250976018)
(4.31303330555084, 1.2413934223704743)
(4.306774176552012, 1.2726477602700077)
(4.299136851401252, 1.3037453837924748)
(4.290128943150362, 1.3346552941504481)
(4.279759431078189, 1.3653466796730651)
(4.268038651739879, 1.3957889465197713)
(4.254978288663175, 1.4259517491769258)
(4.240591360702025, 1.4558050207068698)
(4.224892209059101, 1.4853190027193037)
(4.207896482990183, 1.5144642750350978)
(4.189621124204632, 1.5432117850129665)
(4.1700843499775315, 1.5715328765097727)
(4.14930563499031, 1.5993993184455957)
(4.127305691917956, 1.6267833329450836)
(4.10410645078217, 1.6536576230270454)
(4.079731037091046, 1.679995399814675)
(4.054203748787058, 1.7057704092392854)
(4.027550032026348, 1.730956958210936)
(3.999796455813438, 1.7555299402298639)
(3.9709706855166718, 1.7794648604131882)
(3.941101455290774, 1.8027378599119412)
(3.9102185394340196, 1.8253257396940898)
(3.8783527227085663, 1.8472059836698336)
(3.8455357696535315, 1.868356781136134)
(3.811800392921413, 1.8887570485180958)
(3.7771802206694023, 1.9083864503855354)
(3.7417097630381075, 1.9272254197237775)
(3.705424377751095, 1.9452551774384825)
(3.668360234869541, 1.9624577510750556)
(3.6305542807371274, 1.9788159927339801)
(3.592044201151124, 1.994313596164217)
(3.5528683837963673, 2.008935113017631)
(3.5130658799795773, 2.022665968248238)
(3.4726763657021693, 2.0354924746409315)
(3.4317401021103535, 2.0474018464551915)
(3.3902978953619445, 2.058382212170194)
(3.3483910559499006, 2.068422626318596)
(3.3060613575231232, 2.0775130803972197)
(3.2633509952455753, 2.0856445128437433)
(3.2203025437352255, 2.092808818069464)
(3.176958914624742, 2.0989988545391265)
(3.133363313786244, 2.1042084518897592)
(3.08955919826275, 2.1084324170814264)
(3.0455902329492504, 2.1116665395737635)
(3.0015002470665904, 2.113907595523136)
(2.9573331904715525, 2.115153350996235)
(2.9131330898466805, 2.115402564196913)
(2.868944004813525, 2.1146549867040325)
(2.824809984013057, 2.1129113637190957)
(2.7807750211970204, 2.1101734333234137)
(2.736883011374008, 2.106443924745549)
(2.69317770705396, 2.1017265556407594)
(2.64970267463471, 2.096026028385161)
(2.606501250974053, 2.089348025388295)
(2.5636165001906237, 2.0816992034287787)
(2.521091170736646, 2.0730871870186864)
(2.4789676527853457, 2.0635205608032665)
(2.4372879359755037, 2.053008861003582)
(2.3960935675552726, 2.0415625659105983)
(2.3554256109669747, 2.029193085440194)
(2.315324604914169, 2.0159127497595084)
(2.275830522951792, 2.0017347969959656)
(2.236982733639648, 1.9866733600412165)
(2.198819961298971, 1.9707434524631688)
(2.1613802474111803, 1.9539609535401343)
(2.1247009126973064, 1.9363425924320194)
(2.088818519915881, 1.9179059315043347)
(2.053768837416386, 1.8986693488216475)
(2.019586803484586, 1.8786520198279255)
(1.9863064915152817, 1.85787389823204)
(1.9539610760472146, 1.8363556961174732)
(1.9225827996939593, 1.8141188632960634)
(1.8922029410037886, 1.7911855659263658)
(1.8628517832805356, 1.7675786644179436)
(1.834558584396537, 1.743321690643615)
(1.807351547627752, 1.7184388244823714)
(1.7812577935401204, 1.6929548697163455)
(1.7563033329551947, 1.6668952293058643)
(1.7325130410219867, 1.6402858800672253)
(1.7099106324208786, 1.6131533467784402)
(1.6885186377243133, 1.5855246757387598)
(1.6683583809378302, 1.557427407808336)
(1.649449958243832, 1.528889550954891)
(1.6318122179692742, 1.4999395523347678)
(1.6154627417972416, 1.4706062699361855)
(1.6004178272411451, 1.4409189438129686)
(1.5866924713990063, 1.410907166937427)
(1.5743003560040227, 1.3806008557014355)
(1.5632538337863193, 1.3500302200951257)
(1.5535639161594788, 1.3192257335929092)
(1.5452402622441246, 1.2882181027768567)
(1.5382911692394992, 1.2570382367277084)
(1.5327235641526362, 1.225717216214032)
(1.5285429968933686, 1.1942862627102335)
(1.525753634742058, 1.1627767072743151)
(1.5243582581955588, 1.131219959316394)
(1.5243582581955588, 1.0996474752891199)
(1.525753634742058, 1.0680907273311988)
(1.5285429968933686, 1.0365811718952804)
(1.532723564152636, 1.005150218391482)
(1.538291169239499, 0.9738291978778055)
(1.5452402622441244, 0.9426493318286573)
(1.5535639161594788, 0.9116417010126048)
(1.5632538337863193, 0.8808372145103882)
(1.5743003560040225, 0.8502665789040784)
(1.5866924713990063, 0.819960267668087)
(1.600417827241145, 0.7899484907925455)
(1.6154627417972414, 0.7602611646693285)
(1.631812217969274, 0.7309278822707461)
(1.649449958243832, 0.701977883650623)
(1.66835838093783, 0.673440026797178)
(1.688518637724313, 0.6453427588667542)
(1.7099106324208786, 0.6177140878270737)
(1.7325130410219864, 0.5905815545382885)
(1.7563033329551945, 0.5639722052996495)
(1.7812577935401201, 0.5379125648891684)
(1.8073515476277517, 0.5124286101231426)
(1.8345585843965369, 0.48754574396189876)
(1.8628517832805351, 0.4632887701875703)
(1.8922029410037884, 0.43968186867914816)
(1.922582799693959, 0.41674857130945053)
(1.9539610760472144, 0.39451173848804066)
(1.9863064915152815, 0.372993536373474)
(2.019586803484585, 0.35221541477758866)
(2.053768837416386, 0.33219808578386645)
(2.08881851991588, 0.3129615031011793)
(2.124700912697306, 0.2945248421734944)
(2.1613802474111807, 0.2769064810653794)
(2.19881996129897, 0.2601239821423452)
(2.236982733639648, 0.24419407456429731)
(2.2758305229517912, 0.2291326376095485)
(2.3153246049141685, 0.21495468484600533)
(2.3554256109669747, 0.20167434916532012)
(2.396093567555272, 0.1893048686949158)
(2.4372879359755037, 0.1778585736019318)
(2.478967652785345, 0.16734687380224755)
(2.5210911707366455, 0.1577802475868274)
(2.5636165001906237, 0.14916823117673494)
(2.6065012509740524, 0.14151940921721917)
(2.64970267463471, 0.1348414062203528)
(2.6931777070539593, 0.12914087896475424)
(2.736883011374008, 0.12442350985996475)
(2.780775021197021, 0.12069400128209984)
(2.824809984013056, 0.11795607088641813)
(2.868944004813525, 0.11621244790148122)
(2.9131330898466796, 0.11546487040860054)
(2.9573331904715525, 0.11571408360927893)
(3.001500247066591, 0.11695983908237806)
(3.0455902329492495, 0.11920089503175013)
(3.08955919826275, 0.12243501752408725)
(3.1333633137862438, 0.12665898271575426)
(3.176958914624742, 0.1318685800663869)
(3.220302543735226, 0.13805861653604967)
(3.263350995245575, 0.14522292176177054)
(3.3060613575231232, 0.15335435420829402)
(3.3483910559499, 0.16244480828691743)
(3.3902978953619445, 0.17248522243531983)
(3.4317401021103526, 0.18346558815032177)
(3.472676365702169, 0.19537495996458232)
(3.513065879979577, 0.20820146635727554)
(3.5528683837963664, 0.22193232158788256)
(3.592044201151124, 0.23655383844129652)
(3.6305542807371265, 0.25205144187153317)
(3.6683602348695405, 0.2684096835304579)
(3.705424377751095, 0.2856122571670311)
(3.741709763038107, 0.3036420148817359)
(3.7771802206694023, 0.3224809842199783)
(3.8118003929214126, 0.3421103860874174)
(3.8455357696535315, 0.3625106534693797)
(3.8783527227085663, 0.38366145093568016)
(3.9102185394340196, 0.40554169491142367)
(3.9411014552907737, 0.4281295746935726)
(3.9709706855166713, 0.4514025741923252)
(3.9997964558134376, 0.47533749437564954)
(4.027550032026348, 0.49991047639457786)
(4.054203748787058, 0.5250970253662282)
(4.079731037091046, 0.5508720347908388)
(4.10410645078217, 0.5772098115784678)
(4.127305691917956, 0.60408410166043)
(4.14930563499031, 0.6314681161599183)
(4.170084349977531, 0.6593345580957406)
(4.189621124204632, 0.6876556495925473)
(4.207896482990183, 0.7164031595704154)
(4.224892209059101, 0.7455484318862099)
(4.240591360702025, 0.7750624138986434)
(4.254978288663175, 0.8049156854285877)
(4.268038651739879, 0.8350784880857425)
(4.279759431078189, 0.8655207549324481)
(4.290128943150362, 0.8962121404550654)
(4.299136851401252, 0.9271220508130384)
(4.306774176552012, 0.9582196743355058)
(4.31303330555084, 0.9894740122350394)
(4.317907999161827, 1.0208539095079114)
(4.3213933981843695, 1.0523280859900832)
(4.323486028296922, 1.0838651675379458)
(4.324183803520279, 1.1154337173027566)
};
\addlegendentry{$E_2$}
\end{axis}

\end{tikzpicture}