### 2019-05-05 quali/quali.presentation/figures/cover.tex
WIP

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.7,xscale=0.7]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

% Text Node
\draw  [fill={rgb, 255:red, 126; green, 211; blue, 33 }  ,fill opacity=1 ]  (93, 74) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (93,74) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 219; green, 24; blue, 24 }  ,fill opacity=1 ]  (46, 132) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (46,132) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 126; green, 211; blue, 33 }  ,fill opacity=1 ]  (192, 105) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (192,105) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 245; green, 166; blue, 35 }  ,fill opacity=1 ]  (119, 148) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (119,148) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 245; green, 166; blue, 35 }  ,fill opacity=1 ]  (188, 178) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (188,178) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 208; green, 2; blue, 27 }  ,fill opacity=1 ]  (87, 196) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (87,196) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ]  (155, 227) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (155,227) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 126; green, 211; blue, 33 }  ,fill opacity=1 ]  (281, 75) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (281,75) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 208; green, 2; blue, 27 }  ,fill opacity=1 ]  (234, 132) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (234,132) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 126; green, 211; blue, 33 }  ,fill opacity=1 ]  (380, 105) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (380,105) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 245; green, 166; blue, 35 }  ,fill opacity=1 ]  (307, 148) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (307,148) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 245; green, 166; blue, 35 }  ,fill opacity=1 ]  (376, 178) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (376,178) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 208; green, 2; blue, 27 }  ,fill opacity=1 ]  (275, 196) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (275,196) node [scale=0.8]  {$$};
% Text Node
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (343, 227) circle [x radius= 12.75, y radius= 12.75]   ;
\draw (343,227) node [scale=0.8]  {$$};
% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (54.03,122.1) -- (84.97,83.9) ;

% Connection
\draw [color={rgb, 255:red, 126; green, 211; blue, 33 }  ,draw opacity=1 ][line width=1.5]    (179.83,101.19) -- (105.17,77.81) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (129.99,141.53) -- (181.01,111.47) ;

% Connection
\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (80.12,185.26) -- (52.88,142.74) ;

% Connection
\draw [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ][line width=1.5]    (98.6,201.29) -- (143.4,221.71) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (162.12,216.43) -- (180.88,188.57) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (94.07,185.39) -- (111.93,158.61) ;

% Connection
\draw [color={rgb, 255:red, 245; green, 166; blue, 35 }  ,draw opacity=1 ][line width=1.5]    (176.31,172.92) -- (130.69,153.08) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (242.11,122.16) -- (272.89,84.84) ;

% Connection
\draw [color={rgb, 255:red, 126; green, 211; blue, 33 }  ,draw opacity=1 ][line width=1.5]    (367.8,101.3) -- (293.2,78.7) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (317.99,141.53) -- (369.01,111.47) ;

% Connection
\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (268.12,185.26) -- (240.88,142.74) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (286.6,201.29) -- (331.4,221.71) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (350.12,216.43) -- (368.88,188.57) ;

% Connection
\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=1.5]    (282.07,185.39) -- (299.93,158.61) ;

% Connection
\draw [color={rgb, 255:red, 245; green, 166; blue, 35 }  ,draw opacity=1 ][line width=1.5]    (364.31,172.92) -- (318.69,153.08) ;

\end{tikzpicture}