### 2019-05-31 drafts/3points-ellipse5.tex
wip

\documentclass{article}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathrsfs}

\usepackage{tikz}
\usetikzlibrary{quotes, angles, intersections}
\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\Pp}{\mathscr{P}}
\newcommand{\Cc}{\mathscr{C}}
\newcommand{\E}{\mathscr{E}}

\newcommand{\bigO}{\mathscr{O}}
%\usetikzlibrary{matrix}
%\usepackage{pgfplots}
%\pgfplotsset{compat=1.4}
\usepackage{algorithm2e}
\usepackage{algorithmic}

\usepackage{caption}
\newcommand{\source}[1]{\caption*{Source: {#1}} }

\begin{document}
	\section{Intro}
	Let's assume that $x_1, y_1 = (0,0)$ and $x_2 = 0$. Then we have:

	\begin{align*}
	E=-Cy_2\\
	Dx_3=-(Ax_3^2 + Bx_3y_3 + C(y_3^2-y_2y_3))\\
	\end{align*}

	Renaming things:

	\begin{align}
	E=C\beta\\
	D = A\alpha_a + B\alpha_b + C\alpha_c
	\end{align}

	\section{Working with $a$ and $b$}

	Let's call $q$ the normalizor factor:

	\begin{align*}
	q = \frac{64(BDE - AE^2 - CD^2)}{(4AC-B^2)^2}
	\end{align*}

\end{document}