### 2019-08-08 defesa/tex/chapter/ellipses_n.tex
wip

This section introduces the covering problem that does not have any axis-parallel constraint for the ellipses, we denote this version of the problem as $MCER$. The removal of this constraint introduces a new variable which is responsible for determining the rotation angle of every ellipse.

An instance of the non-axis-parallel is defined exactly like the axis-parallel one on \autoref{chapter:ellipses}. It is given by a set of demand points $\Pp=\{p_1, \dots, p_n\}$, with every point having a unitary weight, and a set of ellipses $\E=\{E_1, \dots, E_m\}$, with fixed shape parameters $(a_i, b_i) \in \R^2_{>0}$, $i \in \{1, \dots, n\}$ with the additional condition that $a_i > b_i$. Given an instance of $MCER$, we define $Q:=(q_1, \dots, q_m) \in \R^{2m}$ to be the centers of each ellipse, $\Theta:=(\theta_1, \dots, \theta_m) \in [0, 2\pi]^m$ to be the angle of rotation of each ellipse and $E_i(q_i, \theta_i)$ to be the coverage region of ellipse $E_i$ with its center at point $q_i$ rotated by angle $\theta_i$. Also, we denote $\Pp \cap E_i(q_i, \theta_i)$ as the set of points covered by $E_i$ on this configuration. Therefore $MCER$ is defined as the problem of determining $Q$ and $\Theta$ (placing and rotating each ellipse) to maximize the number of points covered by the $m$ ellipses, which is given by \autoref{eq:optMCEn}.

\begin{equation}\label{eq:optMCEn}
\max_{Q, \Theta}{\left|\bigcup_{i=1}^{m} \Pp \cap E_i(q_i, \theta_i)\right|}.
\end{equation}

An additional notation is used on this chapter, we define $\tilde{E_i}(q_i, \theta_i)$ to be the set of points on the border of $E_i(q_i, \theta_i)$, we will specially use the operation $\Pp \cap \tilde{E_i}(q_i, \theta_i)$ referring to the set of points from $\Pp$ that lie on the border of $E_i(q_i, \theta_i)$.

\begin{lema}\label{lema:mce_2b}
	Let $(\Pp, \E)$ be an instance of $MCER$. In an optimal solution of $MCER$, for any $E_j \in \E$, such that $|\Pp \cap E_j(q_j, \theta_j)|\ge2$, there is $q_j'$ such that $\Pp \cap E_j(q_j', \theta_j)=\Pp \cap E_j(q_j, \theta_j)$ and $\Pp \cap \tilde{E_j}(q_j', \theta_j) \ge 2$.
\end{lema}

\begin{demonstracao}
	First, the angle of rotation can be ignored as it does not change.

	Let $A=\Pp \cap E_j(q_j, \theta_j)$ be the set of points covered by $E_j$ and $X=\cap_{p \in A}E_j(p, \theta_j)$ be the region of intersection of ellipses centered at each point from $A$.

	As it was shown on \autoref{chapter:ellipses}, $X$ is a region that is limited by arcs of ellipses. As this region is the non-empty intersection of more than one ellipse, there are at least two of these arcs that encounter at one point creating a vertice. Selecting any of these vertices as $q_j'$ will make $|\Pp \cap \tilde{E}_j(q_j', \theta_j)| \ge 2$.

\end{demonstracao}

\begin{definicao}\label{def:feasible_angle}
	Let $E$ be an ellipse and $u, v \in \R^2$. An angle $\theta \in [0, \pi]$ is said to be $(E, u, v)$-feasible if there is $q \in \R^2$ such that $\{u, v\} \subset \tilde{E}(q, \theta)$.
\end{definicao}

\begin{lema}\label{lema:3pnts}
	Let $(\Pp, \E)$ be an instance of $MCER$, in an optimal solution, for any $E_j \in \E$, such that $|\Pp \cap E_j(q_j, \theta_j)|>2$, one of the two cases is true:

	\begin{enumerate}
		\item There is $q', \theta'$, and $\{u, v, w\} \subset \Pp \cap E_j(q_j, \theta_j)$, such that $\{u, v, w\} \subset \tilde{E}(q', \theta')$.

		\item Let $A=\Pp \cap E_j(q_j, \theta_j)$, and $u, v \in A$ such that there exists $\hat{q}_j$ such that $\{u, v\} \subset \tilde{E_j}(\hat{q}_j, \theta_j)$ and $A \subset E_j(\hat{q}_j, \theta_j)$. Then for any $(E_j, u, v)$-feasible angle $\theta \in [0, 2\pi]$, there exists $\bar{q}_j$ such that $\{u, v\} \subset \tilde{E_j}(\bar{q}_j, \theta)$ and $A \subset E_j(\bar{q}_j, \theta)$.
	\end{enumerate}
\end{lema}

The first case of \autoref{lema:3pnts} is saying that there is another optimal solution which has $E_j$ covering the same set of points, but with three points on its border.

The second case of \autoref{lema:3pnts} says that after fixing a pair of points on the border of $E_j$ maintaining the covered set, for any angle that allows the two points to stay on the border of $E_j$, there is a center that maintains the covered set the same.

The idea to prove \autoref{lema:3pnts} is that after fixing $u, v$ on the border of $E_j$, which is possible by \autoref{lema:mce_2b}, the movement of rotation and translation while keeping $u, v$ on the border is continuous. Because of that, the negation of case two implies case one and vice versa.

If we define an equivalence relation between optimal solutions as: $S_1$ is equivalent to $S_2$ if they both cover the same set of points, we can use \autoref{lema:3pnts} and \autoref{lema:mce_2b} to identify the equivalence classes. Let $S$ be any optimal solution, in $[S]$ (its equivalence class) there is another solution where any ellipse $E_j \in \E$ falls in at least one of the cases below:

\begin{itemize}
	\item $E_j$ covers only on point.
	\item $E_j$ covers more than one point with $u$ and $v$ being on the border of $E_j$. Note that there could be infinitely many of solutions like that, however \autoref{lema:3pnts} guarantees that any $(E_j, u, v)$-feasible angle yields an equivalent optimal solution.
	\item $E_j$ covers more than two points with three of them on its border.
\end{itemize}

The following two sections will treat the second and third cases (the first case is trivial). Going through every possibility of an ellipse falling in any of the three cases guarantees that an optimal solution is found.

\section{Ellipse by two points}

Let $E$ be an ellipse with shape parameters $(a, b)\in \R^2_{>0}$ and $u, v\in \R^2$, one wants to find a $(E, u, v)$-feasible angle $\theta\in[0,\pi]$ and every center $q\in\R^2$ such that $\{u, v\} \subset \tilde{E}(q, \theta)$.

For a fixed angle, finding every center such that two points are on the border of the ellipse is done on \autoref{chapter:ellipses_intersection}, from there we know that there could be at most $2$ of such centers.