### 2019-10-16 ONPCE/ellipse_figure.tex
wip

%\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Shape: Ellipse [id:dp1559950964552308] 
\draw   (100,180) .. controls (100,147.42) and (161.34,121) .. (237,121) .. controls (312.66,121) and (374,147.42) .. (374,180) .. controls (374,212.58) and (312.66,239) .. (237,239) .. controls (161.34,239) and (100,212.58) .. (100,180) -- cycle ;
%Straight Lines [id:da3716259356733107] 
\draw    (100,180) -- (374,180) ;

%Straight Lines [id:da9880464900454329] 
\draw    (237,121) -- (237,239) ;

%Shape: Brace [id:dp3973633345998604] 
\draw   (242,182) .. controls (242,186.67) and (244.33,189) .. (249,189) -- (293.5,189) .. controls (300.17,189) and (303.5,191.33) .. (303.5,196) .. controls (303.5,191.33) and (306.83,189) .. (313.5,189)(310.5,189) -- (358,189) .. controls (362.67,189) and (365,186.67) .. (365,182) ;
%Shape: Brace [id:dp94498526835576] 
\draw   (232,129) .. controls (227.34,128.79) and (224.9,131.01) .. (224.68,135.67) -- (224.47,140.23) .. controls (224.16,146.89) and (221.68,150.11) .. (217.01,149.9) .. controls (221.68,150.11) and (223.85,153.55) .. (223.54,160.21)(223.68,157.21) -- (223.33,164.68) .. controls (223.12,169.35) and (225.34,171.79) .. (230,172) ;

%Shape: Circle [id:dp8908034615999807]

% Text Node
\draw (304,205) node   {$a$};
% Text Node
\draw (207,148.33) node   {$b$};
% Text Node

\end{tikzpicture}