### 2019-12-06 defesa/tex/chapter/definicoes.tex
wip

Some definitions and results that are used throughout the text are given in this chapter.

\section{Elliptical and Euclidean norm functions}

A norm in $\R^2$ is a function that maps every vector onto a non-negative real number satisfying homogeneity and the triangle inequality.

Let $u \in \R^2$ be a vector, the Euclidean norm of $u$ is defined as

\begin{equation}\label{eq:norm2}
||u||_2 = \sqrt{u^{T}u}.
\end{equation}

\noindent The elliptical norm, also known as weighted norm, takes a $2$ by $2$ positive definite matrix as its parameter. This matrix can be seen as a linear transformation of the Euclidean norm. The elliptical norm of $u \in\R^2$ is defined as

\begin{equation}
||u||_{Q} = \sqrt{u^{T}Qu},
\end{equation}

\noindent where $Q$ is a $2$ by $2$ positive definite matrix.

The elliptical norm, when taking $Q$ to be the identity matrix, becomes the Euclidean norm.

Determining the distance between two points, given a norm function, is done by calculating the norm of the vector defined by the difference between the two points. For example, the elliptical distance between the points $p,q \in \R^2$ is given by $||p-q||_{Q}$.

\section{Disk}

A circle (or circumference) is a set of points in $\R^2$ that have the same Euclidean distance, also known as radius, to another point, also referred to as the center of the circle. A unit circle is a circle with radius equal to $1$.

A disk is the set of points bounded by a circle. In other words let $c \in \R^2$. A unit disk with center $c$ is the set of every point $p \in \R^2$ which satisfies

\begin{equation}\label{eq:disk}
||p-c||_2^2 \le 1.
\end{equation}

\section{Ellipse}

An ellipse is a curve which is categorized, along with the parabola and the hyperbola, as a conic section. As the name suggests, conic sections are curves resulted from the intersection of a right circular cone in $\R^3$ with a plane \cite{brannan:geometry}. From that definition, an equation which describes any conic section is given as follows

\begin{equation}\label{equation:quadratic_form}
Ax^2 + Bxy + Cy^2 + Dx + Ey + F = 0,
\end{equation}

\noindent where $A,B,C,D,E,F \in \R$ are fixed and $x, y \in \R$.

To distinguish an ellipse from the other conic sections given an instance of \autoref{equation:quadratic_form}, the condition $4AC - B^2>0$ must be verified \cite{ayoub}.

Assuming the center of an ellipse is $c \in \R^2$, then \autoref{equation:quadratic_form} can be rewritten as a quadratic form as follows

\begin{equation}
(p-c)^{T}Q(p-c) = 1,
\end{equation}

\noindent with $p \in \R^2$ and $Q$ being a $2$ by $2$ positive definite matrix which carries the parameters of the ellipse. From \autoref{equation:quadratic_form}, $Q$ can be defined as follows

\[
Q=
\left( {\begin{array}{cc}
	A & \frac{B}{2} \\
	\frac{B}{2} & C \\
	\end{array} } \right).
\]

Note that asking $Q$ to be positive definite is the same as asking $4AC-B^2$ to be positive. This makes us arrive at the following definition of the ellipse.

\begin{definicao}\label{def:ellipse}
    Let $c\in \R^2$ be the center of an ellipse and $Q$ be a $2$ by $2$ positive definite matrix. An ellipse is the set of every point $p \in \R^2$ such that $||p-c||_{Q}^2 = (p-c)^{T}Q(p-c) = 1$. Also, a point $p$ is considered covered by an ellipse if $||p-c||_{Q}^2 = (p-c)^{T}Q(p-c) \le 1$.
\end{definicao}

An alternative way to define an ellipse, which can be seen as just a property derived from the definition above, is to begin its construction with two points called foci and a constant $R \in \R$, with $R$ being greater than the Euclidean distance between the two foci points (see \autoref{fig:ellipse_with_foci}). The ellipse is, then, defined as the set of points whose distance to the foci is equal to $R$. In other words, let $f_1, f_2 \in \R^2$ be the two foci points, the ellipse is the set of every point $p \in \R^2$, such that $||p-f_1||_2 + ||p-f_2||_2 = R$. It can be shown that this definition is equivalent to \autoref{def:ellipse}, with the coverage of a point $p$ being equivalent to $||p-f_1||_2 + ||p-f_2||_2 \le R$.

\begin{figure}[H]
    \centering

    \caption{A non-axis-parallel ellipse and its foci points.}
    \input{tex/figures/ellipse_with_foci.tex}
    \fautor
    \label{fig:ellipse_with_foci}
\end{figure}

Also, in \autoref{fig:ellipse_with_foci}, the distance $a = ||p_a - c||_2$, where $p_a$ is one of the intersection points of the ellipse with the major axis, is called the semi-major, and the distance $b = ||p_b-c||_2$, where $p_b$ is one of the intersection points of the ellipse with the minor axis, is called the semi-minor. These two values are also referred to as the shape parameters of an ellipse. Let $d = ||c-f_1||_2$, then it is easy to see that $a = R - d$ and $b = \sqrt{\frac{R^2}{4} - d^2}$.

Finally, an ellipse is said to be axis-parallel if its major-axis (see \autoref{fig:ellipse_with_foci}), which is the line that passes through its two foci points, is parallel to the $x$-axis.

\subsection{Axis-parallel}

An axis-parallel ellipse centered at $c = (c_x,c_y)$ can be described using \autoref{def:ellipse} with $Q$ being a diagonal matrix \footnote{The only non-zero terms are in the main diagonal.}. This can be understood as a scaling transformation applied to the Euclidean norm.

Defining the matrix $Q$ as

\[
Q=
\left( {\begin{array}{cc}
    \frac{1}{a^2} & 0 \\
    0 & \frac{1}{b^2} \\
    \end{array} } \right),
\]

\noindent then, starting from \autoref{def:ellipse}, we can obtain the following equation

\begin{align*}
        (p-c)^{T}Q(p-c) = 1 & \Rightarrow \\
    (\frac{p_x-c_x}{a^2}, \frac{p_y-c_y}{b^2})^{T}(p_x-c_x, p_y-c_y) = 1 & \Rightarrow
 \end{align*}
 \begin{equation}\label{equation:pellipse}
  \frac{(p_x-c_x)^2}{a^2} + \frac{(p_y-c_y)^2}{b^2} = 1,
 \end{equation}

\noindent where $a$ and $b$ are the semi-major and semi-minor shape parameters, respectively.

Also, the coverage region is determined by just changing the equality to a inequality as follows

\begin{equation}\label{equation:cover_pellipse}
\frac{(p_x-c_x)^2}{a^2} + \frac{(p_y-c_y)^2}{b^2} \le 1.
\end{equation}

Another way to represent ellipses, which will be useful in some occasions, is through writing it as a curve, function of the angle with its major-axis (see \autoref{fig:ellipse_params}).

\begin{figure}[H]
    \centering

    \caption{The ellipse as a parametric curve.}
    \input{tex/figures/ellipse_definition.tex}
    \fautor
    \label{fig:ellipse_params}
\end{figure}

Let $c \in \R^2$ be the center of an ellipse with shape parameters $(a,b) \in \R^2_{>0}$. Then $\gamma : [0, 2\pi] \mapsto \R^2$ defines a curve which maps every angle onto a point on the ellipse and it is defined as follows

    \begin{equation}\label{eq:parametric_ellipse}
    \gamma(t) = \left\{
    \begin{array}{l}
    x(t)= a\cos{t} + c_x,\\
    y(t)=b\sin{t} + c_y.
    \end{array}
    \right.
    \end{equation}

No equivalent disk-circle wording exists for ellipses, this could be a source of ambiguity in the text, that is why a note for the reader was judged to be necessary. Throughout this work an ellipse will represent the set of points that satisfy \autoref{def:ellipse}. In some places, though, with prior clarification, we will denote as an ellipse the set of points that are covered by the ellipse itself. For example, when we define $\Pp \cap E$ as the set of points in $\Pp$ that are covered by $E$, we are implicitly calling $E$ the set of points that are covered by the ellipse itself as it is defined by \autoref{def:ellipse}.

\subsection{Non-axis-parallel}

An non-axis-parallel ellipse centered at $c\in\R^2$ can also be described by \autoref{def:ellipse}. Nonetheless, a simpler equation is going to be used here. Besides the center, the shape parameters $(a, b) \in \R^2_{>0}$, with $a > b$, an angle of rotation $\theta \in [0, \pi]$ is given. This angle represents the angle between the $x$-axis and the major-axis of the ellipse.

A rotated ellipse by $\theta$ can be transformed into a axis-parallel centered at the origin by applying two linear transformations. First, the ellipse can be translated to the origin and then rotated by $-\theta$, the reverse transformations are expressed by \autoref{eq:rotated_ellipse}.
\begin{equation}\label{eq:rotated_ellipse}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}=1.
\end{equation}
Also, the coverage region of an ellipse centered at $c\in\R^2$ with shape parameters $(a, b)\in\R^2_{>0}$ and rotation angle $\theta \in [0, \pi]$ is represented by every point $(x, y) \in \R^2$ that satisfies \autoref{eq:rotated_ellipse_co}.
\begin{equation}\label{eq:rotated_ellipse_co}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}\le 1.
\end{equation}

\section{Complex numbers}

We introduce here some things about complex numbers that are used later on \autoref{chapter:e3p}.

Any complex number $z = a + bi \in \mathbb{C}$ can be written as follows:
\begin{equation}
z = re^{i\theta},
\end{equation}
where $\theta$ is the polar angle of the vector $(a, b)$, we refer to it as $angle(z)$, and $r$ is the length of this vector, we refer to it as $|z| = \sqrt{a^2 + b^2}$. This is known as the Euler's Formula. If $r=1$, then we say that $z$ is in the unit circle. The complex conjugate is another important operator that we need. Let $z = a + bi \in \mathbb{C}$, then we refer to $\bar{z}$ as the complex conjugate of $z$ and it is defined as:
\begin{equation}
\bar{z} = a - bi.
\end{equation}
Also, it is useful to see that, then $1/z=\bar{z}/|z|$. Two observations about the angles of complex numbers are worth mentioning:
\begin{align}
angle(\bar{z}) = -angle(z)\\
angle(-z) = \pi - angle(z).
\end{align}

\section{Polynomials and their roots}

In this work, only univariate polynomials are used, so the definition and results are focused on this case.

A function $p_n: \mathbb{C} \mapsto \mathbb{C}$ is a $n$-degree polynomial if it can be written as:
\begin{equation}\label{eq:poly}
p_n(z) = \sum_{k=0}^{n} a_k z^k.
\end{equation}
The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula to determine their roots. However, a numerical approach exists for higher-degree polynomials.

 In \citeonline[p.~195]{horn} a theorem is presented which says that for every univariate polynomial, there exists a companion matrix, such that its eigenvalues are the zeros of that polynomial.  For example, the companion matrix of a degree-$5$ polynomial written as \autoref{eq:poly} is given by:
 \begin{equation}
 \left[\begin{array}{ccccc}
 0 & 1 & 0 & 0 & 0\\
 0 & 0 & 1 & 0 & 0\\
 0 & 0 & 0 & 1 & 0\\
 0 & 0 & 0 & 0 & 1\\
 -\dfrac{a_0}{a_5} & -\dfrac{a_1}{a_5} & -\dfrac{a_2}{a_5} & -\dfrac{a_3}{a_5} & -\dfrac{a_4}{a_5}
 \end{array}\right].
 \end{equation}

 Finding every eigenvalue of a matrix can be done using the QR algorithm, which runs in $\bigO(n^3)$ and uses $\bigO(n^2)$ memory. In \cite[p.~341]{golub} a practical version of it is developed. The idea is to convert the input matrix to the Hessenberg form, after that, the matrix preserves its form through the iterations and under some assumptions, convergence to the eigenvalues is achieved. It is worth mentioning that this approach is actually used in MATLAB and LAPACK, which is the external library used in NumPy for Python.
 Also, in \citeonline{barel} it is pointed that companion matrices are already in the Hessenberg form, making the Hessenberg conversion step unnecessary. This and other properties are used in \citeonline{barel} on the development of a $\bigO(n^2)$ algorithm for the specific case of companion matrices.

\section{Real trigonometric polynomial}

The same definition found in \citeonline[p.~150]{powell} for real trigonometric polynomials is given here. They are also referred to as truncated Fourier Series in \citeonline{boyd:2006} and are given by:
\begin{equation}\label{eq:trig_poly}
p_n(\theta) = \sum_{k=0}^{n} a_k\cos(k\theta) + \sum_{k=1}^{n} b_k\sin(k\theta).
\end{equation}
We say that $p_n$ as defined by \autoref{eq:trig_poly} is a $n$-degree real trigonometric polynomial. An important property is stated in \citeonline[p.~150]{powell}, it says that a $n$-degree polynomial can have up to $2n$ distinct roots on the interval $[0, 2\pi)$. It also says that a function written in the format:
$$\begin{array}{cc}\cos^j{\theta}\sin^k{\theta} & \quad j, k \in \mathbb{Z}_+ \end{array},$$
can be transformed into a real trigonometric polynomial of degree $j+k$. Therefore, the following expression also represents a $n$-degree real trigonometric polynomial:
\begin{equation}\label{eq:trig_poly_2}
\sum_{0 \le j+k \le n}c_{j,k}\cos^j{\theta}\sin^k{\theta},
\end{equation}
for some $\{c_{j,k} \in \R: 0 \le j+k \le n\}$.