### 2019-12-30 defesa/tex/chapter/pmclp.tex
wip

In this chapter, we introduce a version of the classical Euclidean norm PMCLP where each facility has a given coverage radius. 
We refer to this problem as \sigla{MCD}{Maximum Covering by Disks}. We also propose a method for it intending to adapt it for the elliptical PMCLP that will be introduced in the next chapter.

\section{Definition}

An instance of MCD is given by a set of $n$ demand points $\Pp:=\{p_1, \dots, p_n\}$, with $p_j\in\R^2$; and a set of $m$ disks $\D:=\{D_1, \dots, D_m\}$, which represents the facilities and are used to cover the demand set. Additionally, for every point in $\Pp$ a weight is given, and for every disk in $\D$ a coverage radius is given. The weight of the $j$-th point is denoted by $w_j\in\R_{>0}$, and the coverage radius of the $j$-th facility is denoted by $r_j \in \R_{>0}$.

A solution for an instance of MCD is determined by $Q:=(q_1, \dots, q_m) \in \R^{2m}$ which specifies the center of every disk in $\D$.
The coverage area of $j$-th disk centered at $q_j$ is denoted by $D_j(q_j)$. Finally, let $w : 2^{\Pp} \mapsto \R_{>0}$ be the function

\begin{equation}
w(A) = \sum_{j : p_j \in A} w_j,
\end{equation}
which takes a subset of $\Pp$ and returns the sum of the weights of every point in that subset; then an optimal solution of MCD is given by the optmization problem:

\begin{equation}
\max_{Q} w\left(\bigcup_{j=1}^m \Pp \cap D_j(q_j)\right).
\end{equation}

It is important to note that MCD is a slightly different PMCLP than the one introduced in the first study on the subject in \citeonline{church:1984}. There, a coverage radius is given for each demand, rather than for each facility, and a demand is only considered covered if a facility is located within its radius.

\subsection{CLS and CIPS}

The method proposed in \citeonline{church:1984} involves the creation of a \sigla{CLS}{Candidate List Set} which is a finite set of possible solutions for each facility. After a CLS is constructed, a complete search can be executed to find the best solution.

As this work is developed towards an exact method for MCD, the focus is on constructing a CLS that contains an optimal solution.
This is done in \citeonline{church:1984} where a circle is defined for every demand point, and a \sigla{CIPS}{circle intersection point set} is constructed containing the demand points and every pairwise circle intersection.
We introduce a similar approach based on works found on the case for only one disk and describe a $\bigO(n^2\lg{n})$ algorithm for it.

\section{One disk version}

This version of the problem will be refered to as \sigla{MCD-1}{Maximum Covering by One Unit Disk} and it is just a specific case of MCD with one unit disk ($m=1$ and $r_1=1$). We later use the algorithm for MCD-1 to construct a CLS which is guaranteed to contain an optimal solution for MCD.

Two exact methods for it have been found in the literature. A $\bigO(n^2)$ algorithm is proposed by \citeonline{chazelle:1986} which improved the previously $\bigO(n^2\log{n})$ one proposed by \citeonline{drezner}.
As it has been mentioned, MCD-1 is a 3SUM-HARD problem, which means that it is as hard as the 3SUM problem (the problem of finding $3$ real numbers that sum to $0$, given $n$ real numbers). Initially the lower bound of the 3SUM problem was conjectured to be $\Omega(n^2)$, matching the best algorithm for MCD-1, which meant that no better time-complexity could be achieved. Since then, however, better algorithms for 3SUM have been developed with a $\bigO(\frac{n^2}{poly(n)})$ run time complexity \cite{3SUM-kopelowitz:2014}.

In \citeonline{drezner}, the main idea used to develop the $\bigO(n^2\log{n})$ algorithm is that, even though there are infinitely many points where the disk could be placed, only a few of them, a finite amount of $\bigO(n^2)$, needs to be considered for the method to find an optimal one.
The algorithm, for every point, sorts the other points with respect to the angle they form with the first one. After that, the first point is placed on the border of the disk and, going through the sorted list, the algorithm inserts and removes points from the disk coverage. Also, when inserting and removing a point from the coverage, it only checks the disk centers that make the entering/leaving point to be on the border. Because the algorithm only checks the centers that make the disk have two points on its border, the number of centers it goes through is bounded by the number of pairs of points, which is $\binom{n}{2} = \bigO(n^2)$.

In \citeonline{chazelle:1986} and \citeonline{cabello:2006}, on the other hand, instead of working directly with MCD-1, an equivalent problem called \sigla{MWC}{Maximum Weight Clique} was introduced. The algorithm for MCD that is developed in this chapter also uses that equivalence. Because of that, MWC is introduced in the next section.

\section{Maximum Weight Clique Problem}

An instance of MWC is given by a set of points $\Pp:=\{p_1,\dots,p_n\}$, with $p_i \in \R^2$; and a set of disks $\D:=\{D_1,\dots,D_n\}$, such that $D_i$ is centered at $p_i$, for $i=1,\dots,n$, and with the $i$-th disk having a weight $w_i\in \R_{>0}$ assigned to it.

A clique, in this context, is a non-empty intersection region of one or more disks, and its weight is the sum of the weights of those disks in the intersection.
Following this, a solution for MWC can be defined as just a point $q\in\cup_{j=1}^n D_j$ as from $q$, the corresponding clique $S$ can be obtained as follows:

\begin{equation}
S = \bigcap_{j : q \in D_j} D_j.
\end{equation}
With a geometric observation, the number of possible values for the solution can be reduced.
Let $\partial \D = \{ \partial D_1, \dots, \partial D_n \}$ be the set of circles corresponding to each disk in $\D$, it is possible to see that any clique has at least one point on its border that is in the intersection of two circles in $\partial D$ or, in the case where the clique is composed of only one disk, the clique contains obviously the center of that disk. Therefore, $q$ can be actually limited to the set of pairwise intersections of $\partial \D$ as well as $\Pp$. Finally, an optimal solution of MWC is given by the optimization problem

\begin{equation}
\max_{q} \sum_{D_k \cap q \neq \emptyset} w_k,
\end{equation}
with $q \in \{\partial D_i \cap \partial D_j : 1 \le i < j \le n\} \cup \Pp$. It can be shown that the number of possible solutions for an instance of MWC is bounded by $\binom{n}{2} + n = \bigO(n^2)$.

It is worth pointing out that MWC is a different problem than the maximum clique on a intersection graph (a graph where the vertices are the disks and an edge exists if there is an intersection between two disks). 
As shown in \autoref{fig:three_disks_no_intersection}, three disks could have non-empty pairwise intersection and still have an empty intersection of all of them together.
That is why MWC is also referred to as the Maximum Geometric Clique Problem and the other version, when there is only the pairwise intersection condition, is referred to as the Maximum Graphical Clique Problem \cite{inplace:2014}.

%The graphical version of the problem was studied by \citeonline{graphical-clique}, where a $\bigO(n^{4.5})$ algorithm was proposed. Also, in \citeonline{inplace:2014}, a $\bigO(n^2\log{n})$ time in-place algorithm\footnote{An in-place algorithm is an algorithm that needs $\bigO(1)$ extra space.} for arbitrary radii disks was proposed.

In \citeonline{chazelle:1986}, the method for MWC consists on building a planar graph on which the vertices were the $\binom{n}{2}$ pairwise intersection of the circumferences and the edges were the arcs of the circumferences connecting the intersections. With the graph constructed, a traversal is done to obtain the answer, thus the time complexity of $\bigO(n^2)$.

\begin{figure}[H]
\centering

    \caption{Three disks that have non-empty pairwise intersection among them, but no common intersection.}
    \begin{tikzpicture}[scale=1]
    %\draw [help lines] (-5,-3) grid (5,3);
    \draw (0,0) circle (1cm) (0.9,1.6) circle (1cm) (-0.9,1.6) circle (1cm);

    %a^2-b^2=c^2 -> c^2=25-9=16 -> c=4

    \draw[fill] (0.9,1.6) circle [radius=.5pt];
	\draw[fill] (0,0) circle [radius=.5pt];
    \draw[fill] (-0.9,1.6) circle [radius=.5pt];

    \node[below] at (0,0) {$p_1$};
    \node[below] at (0.9, 1.6) {$p_2$};
    \node[below] at (-0.9, 1.6) {$p_3$};

    %\draw [-] (-5,0) -- (5,0);
     %\draw [-] (0,-3) -- (0,3);
     %\draw [|-|] (0.001,-0.1) -- (4.999,-0.1);
    \end{tikzpicture}
    \label{fig:three_disks_no_intersection}
    \fautor
\end{figure}

As it has been mentioned, with the equivalence of the two problems, an optimal solution of the Maximum Weight Clique Problem is also an optimal solution of the MCD-1, which means that a disk centered at $q^*$ which is an optimal solution of MWC, will have a maximal weight covering of the demand set $\Pp$.

Given an instance of MCD-1, the equivalent MWC intance is obtained by defining set $\D$ to contain the disks centered at $\Pp$ and setting the weight of every disk to be the weight of its corresponding point in $\Pp$. A disk $D_i$ will represent the area where a disk can be placed in order to cover $p_i$. This means that a intersection between some disks is a region where a disk could be placed to cover the corresponding points.

In \autoref{fig:three_disks_no_intersection}, it can be seen that there is no point where a disk could be placed such that it would cover $p_1, p_2$ and $p_3$, nonetheless, in any of the pairwise intersections, a disk could be placed to cover the two corresponding points.

Formally, in MWC, if a point $q$ lies inside $\bigcap_{k \in I} D_k$, with $I \subset \{1,\dots,n\}$, then a disk centered at $q$ will cover the points $p_k$, with $k\in I$ in the MCD-1 problem. Conversely, the same applies for a disk placed at $q$ that covers points $p_k$, with $k \in I$ in the one disk maximal covering problem. It means that $q$ will lie inside region $\bigcap_{k \in I} D_k$.

\subsection{An algorithm for the Maximum Weight Clique Problem}

The algorithm described here is based on the one in \citeonline{drezner}, also with some ideas from \citeonline{inplace:2014} and \citeonline{cabello:2006}. It has a run time complexity of $\bigO(n^2\log{n})$ and uses $\bigO(n)$ of extra space. It is worth noting, however, that a $\bigO((n+K)\log{n})$ run time, with $K$ being the number of intersections, can be obtained by using the algorithm in \citeonline{bentley:1979} to find all the intersections among the $n$ circumferences.

Without loss of generality, the weights will be ignored, and the method will be described for MWC, assuming that every disk has unit weight. Also, it will be assumed that no pair of disks are placed at the same center.

Let $\D$ be a set of $n$ unit disks, a non-empty intersection area of a subset of disks is convex and bounded by the arcs of the disks that are intersecting \cite{inplace:2014}.
With this observation, in order to find an optimal solution, it is sufficient to check, for every disk $D$, every intersection area that is bounded by its arc.

\begin{definicao}\label{def:inter_arc}
    Let $D_i$ and $D_j$ be two unit disks that intersect (at least at one point). Also let $(\theta_1, \theta_2) \in [0,2\pi]^2$ be the two angles that the circumferences induced by $D_i$ and $D_j$ intersect, with the condition that $(\theta_1,\theta_2)$ defines an arc (counter-clockwise order) of $D_i$ that is the border of $D_i \cap D_j$. If $D_i$ is tangent to $D_j$, then $\theta_1=\theta_2$. Then, define $\Gamma_+(i,j) = \theta_1$ and $\Gamma_-(i,j) = \theta_2$, also we refer to them as opening and closing intersection angles respectively.
\end{definicao}

\begin{figure}[H]
\centering

    \caption{Three disks and their intersection points.}
    \input{tex/img2.tex}
    \fautor
    \label{fig:3disks_intersect}
\end{figure}

In \autoref{fig:3disks_intersect}, it is shown all the intersection points between $D_1$ with $D_2$ and $D_3$. Also, they are labeled according to \autoref{def:inter_arc}. Note that $\Gamma_+(1,3) > \Gamma_-(1,3)$ (the angles should be in the $[0,2\pi]$ interval).

With \autoref{def:inter_arc} in hand, we can establish the basis of the algorithm to find the maximum clique in which a disk $D_i$ participates: a traversal going through every point of intersection with $D_i$, in counter-clockwise order, keeping a set of active disks. When an opening intersection angle is reached, the corresponding disk is added to the active set; when a closing one is reached, the corresponding disk is removed from the active set.
This simple traversal, however, would not handle the special case with $\Gamma_+(i,j) > \Gamma_-(i,j)$ (see \autoref{fig:3disks_intersect}). If the traversal begins at the point with smallest angle, the algorithm would remove $D_3$ from the active disks without first adding it. If there was another disk starting before $\Gamma_-(1,3)$, the algorithm would not have both of them in the active set at the same time, and an optimal solution could end up not being found. This can be worked around repeating the traversal once without resetting the active disks set, that way, in the beginning of the second traversal, the active set would contain the disks that have $\Gamma_+(i,j) > \Gamma_-(i,j)$. This is shown in \autoref{fig:array_disks}, where the intersections list is duplicated, simulating the traversal repetition (note the indication to where the traversal starts as well as the positive and negative signs representing when a intersection with another disk opens and closes, respectively). It can be seen that without the repetition, the algorithm would find $2$ as the value of an optimal solution, which is wrong.

\begin{figure}[H]
    \centering

    \caption{The intersections list of a disk with three other disks.}
    \input{tex/figures/array_disks.tex}
    \fautor
    \label{fig:array_disks}
\end{figure}

\begin{algoritmo}
\caption{Algorithm for $MCD(\Pp, 1)$ with unit weights}\label{algoritmo:mcd_1}
\begin{algorithmic}[1]
\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$.}
\Ensure{The maximum number of points that can be covered by a unit disk.}

\item[]

\Procedure{$MCD_1$}{$\Pp$}
\State $Q_{best} \gets \{\}$
\State $ans \gets$ center of $D_1$
\ForAll{$p_i \in \Pp$}
\State Let $D_i$ be the disk with center at $p_i$
\State Let $I_i$ be the set of disks that intersect with $D_i$
\State $A \gets \{\}$ \Comment{The multiset of intersection angles with $D_i$.}

\ForAll{$j \in I_i$}
\State $A \gets A \cup \{\Gamma_+(i,j) \cup \Gamma_-(i,j)\}$
\EndFor

%\State $A = \bigcup_{j \in I_i} \Gamma_+(i,j) \cup \Gamma_-(i,j)$
\State $Q \gets \{D_i\}$ \Comment{The set of active disks.}
\For{$cnt=1..2$} \Comment{Do it twice.}
\For{$a \in A$}\Comment{Assuming $A$ is sorted.}
\State Let $D_a$ be the disk that intersects $D_i$ at angle $a$. 
\If{$a$ is a starting angle}
\State $Q \gets Q \cup \{D_a\}$
\Else
\State $Q \gets Q \setminus \{D_a\}$
\EndIf
\If{$|Q_{best}| < |Q|$}
\State $Q_{best} \gets Q$
\State $ans \gets $ point corresponding to the intersection angle $a$
\EndIf
\EndFor
\EndFor
\EndFor

\State \Return $ans$
\EndProcedure
\end{algorithmic}
\end{algoritmo}

\begin{lema}\label{lema:disk}
\autoref{algoritmo:mcd_1} for solving the Maximum Clique Problem has a $\bigO((n+K)\log{n})$ run time complexity, where $K$ is the number of intersections of the $n$ disks.
\end{lema}

\begin{proof}
    Finding every intersection can be done in $\bigO((n+K)\log{n})$  by a plane sweep, the method is described in \cite{bentley:1979}. 
    Because the traversal is made in counter-clockwise order, the intersection points have to be sorted by their intersection angles, so an additional $\bigO(K\log{K})$ pre-processing is needed. All the other operations can be done in constant time. Therefore, the final algorithm complexity is $\bigO((n+K)\log{n})$.
\end{proof}

If a simpler implementation is desired, or the number of intersections is large, determining the set $I_i$ (the set of disks that intersect with $D_i$, defined in \autoref{algoritmo:mcd_1}) can be simply done in $\bigO(n^2)$, making the algorithm have a worst-case complexity of $\bigO(n^2\log{n})$.

\section{Multiple disks, $MCD(\Pp, m)$}

In \cite{cabello:2006}, a $\bigO(n^{2m-1}\log{n})$ algorithm for $MCD(\Pp, m)$ is developed as a sub-routine for its $(1-\epsilon)$-approximation algorithm. Firstly, they solve the sub-problem $MCD(\Pp, 2)$ in $\bigO(n^3\log{n})$. Then for the rest of the points that are not in that solution, it uses the algorithm developed in \cite{chazelle:1986} for the one-disk case, checking every possible solution for every one of the disks left.

Also, in \cite{zhou} an heuristic method for large-scale $MCD$ is proposed. It uses a probabilistic algorithm called mean-shift which is a gradient ascent method proven to converge to a local density maxima of any probability distribution. The mean-shift is utilized to find good candidates of centers for the unit disks, then the method backtracks to find the best assignment. The results showed that the greedy algorithm achieves an optimal coverage in some instances, but for some other ones it has a 15 percent worse coverage ratio.

It can be seen that in any solution of $MCD(\Pp, m)$, a disk placed at a point $q$ that covers at least one point $p \in \Pp$ has a correspondence to the Maximum Weight Clique Problem: the point $q$ is inside an intersection area of at least one disk and that area is bounded by some disk, which means it will be checked by \autoref{algoritmo:mcd_1} as a candidate to be an optimal solution. The number of points \autoref{algoritmo:mcd_1} goes through is $\bigO(n^2)$, then checking every possible center for every ellipse yields a $\bigO(n^{2m})$ run-time complexity.
This algorithm is described in \autoref{chapter:ellipses} for the ellipses case.

The choice of developing a different method for the problem, instead of using the one from \cite{cabello:2006}, is taken for the sake of simplicity, considering both algorithms achieve similar bounds.