### 2020-02-17 defesa/tex/chapter/mcer.tex
wip

This chapter introduces the elliptical PMCLP where there is no axis-parallel constraint, and the ellipses can be freely rotated. We refer to this problem as \sigla{MCER}{Maximum Covering by Ellipses with Rotation}. In comparison with MCE, this problem introduces a new variable that is responsible for determining the rotation angle of every ellipse, making MCER a more challenging problem.

\section{Definition}

An instance of the non-axis-parallel is defined exactly like the axis-parallel one on \autoref{chapter:mce}. It is given by a set of demand points $\Pp=\{p_1, \dots, p_n\}$, $p_j\in\R^2$; a list of weights $\Ww:=\{w_1, \dots, w_n\}$, with $w_j\in\R_{\ge0}$ being the weight of point $p_j$;
and $m$ ellipses given by their shape parameters $\Rr:=\{(a_1, b_1), \dots, (a_m, b_m)\}$, with $(a_j, b_j)\in\R_{>0}^2$ and $a_j>b_j$.
Additionally, to make the text more clear, we define a set of $m$ functions that represent the coverage regions of each ellipse as $\E = \{E_1, \dots, E_m\}$, with $E_j\colon \R^2\times\R \mapsto \R^2$ being a function that takes the center and angle of rotation where the $j$-th ellipse is located as input, and returns its coverage region as defined by \autoref{eq:rotated_ellipse_co}.
Lastly, an instance of MCER is defined as the tuple $(\Pp, \Ww, \Rr)$.

Given an instance of $MCER$, we define $Q:=(q_1, \dots, q_m) \in \R^{2m}$ as the centers of each ellipse, and $\Theta:=(\theta_1, \dots, \theta_m) \in [0, \pi)^m$ as the angles of rotation of each ellipse. Then, we define MCER as the problem of determining $Q$ and $\Theta$ (placing and rotating each ellipse) to maximize the weight of the points covered by the $m$ ellipses given by
\begin{equation}\label{eq:optMCEn}
\max_{Q, \Theta}{w\left(\bigcup_{i=1}^{m} \Pp \cap E_i(q_i, \theta_i)\right)}.
\end{equation}
In addition to that, we define an equivalence relation between solutions of MCER. We say that two solutions are equivalent if the set of points covered by them is the same. That is, two solutions of MCER $(Q, \Theta)$ and $(Q', \Theta')$ are said to be equivalent if, and only if 
$$\bigcup_{j=1}^m\Pp \cap E_j(q_j', \theta_j') = \bigcup_{j=1}^m\Pp \cap E_j(q_j, \theta_j).$$

In the next section we present some results which ultimately lead up to the construction of a finite set that contains at least one optimal solution for MCER.

\section{An optimal and finite set of solutions}

In this section, we construct a finite list of centers and angles of rotation, also referred to as a \sigla{CLS}{Candidate Locations Set}, for each ellipse and show that at least one optimal solution is in the set of solutions created from those lists. The results presented in this section are strongly based on \autoref{chapter:e3p}, more specifically on \autoref{lema:e3p}, which states that there exists at most six solutions for any instance of E3P.

We start by introducing a lemma, which says that given an optimal solution of MCER, it is always possible to find an equivalent one, such that every ellipse covering more than one point contains two of them.

\begin{lema}\label{lema:mce_2b}
	Let $(Q^*, \Theta^*)$ be an optimal solution of an instance $(\Pp, \Ww, \Rr)$ of MCER. 
	Then, for any $j\in\{1, \dots, m\}$ with $|\Pp \cap E_j(q_j^*, \theta_j^*)|\ge2$, 
	an equivalent solution $(Q', \Theta^*)$ exists, such that $|\Pp \cap \partial E_j(q_j', \theta_j^*)|\ge 2$.
\end{lema}

\begin{proof}
	First, the angle of rotation can be ignored as it does not change.

	Let $A=\Pp \cap E_j(q_j^*, \theta_j)$ be the set of points covered by the $j$-th ellipse and $X=\cap_{p \in A}E_j(p, \theta_j)$ be the region of intersection of ellipses centered at each point in $A$.

	As it was shown on \autoref{chapter:mce}, $X$ is a region that is limited by arcs of ellipses. As this region is the non-empty intersection of more than one ellipse, there are at least two of these arcs that encounter at one point, creating a vertex. Selecting any of these vertices as $q_j'$ will make $|\Pp \cap \partial E_j(q_j', \theta_j)| \ge 2$.

\end{proof}

What \autoref{lema:mce_2b} states is that transforming any optimal solution of MCER into an equivalent optimal solution where every ellipse that covers more than one point contains two points is possible (an example can be seen in \autoref{fig:ellipse-2-points}). \autoref{lema:mce_2b} also states that this equivalent optimal solution can always be achieved by just translating the ellipses; that is, no change in the angle of rotation is required.

\begin{figure}
	\centering
	\caption{An optimal solution before and after applying \autoref{lema:mce_2b}.}
	\includegraphics[scale=.28]{tex/figures/ellipse-2-points}
	\fautor
	\label{fig:ellipse-2-points}
\end{figure}

Next, we define, for an optimal solution, a set of equivalent solutions, such that any ellipse covering more than one point contains at least two points.

\begin{definicao}
	Let $(Q^*, \Theta^*)$ be an optimal solution for an instance $(\Pp, \Ww, \Rr)$ of MCER. We define $\Pi(Q^*, \Theta^*)$ as the set of every equivalent solution of $(Q^*, \Theta^*)$, such that for any $(Q, \Theta)\in\Pi(Q^*, \Theta^*)$, for $j\in\{1, \dots, m\}$ with $|\Pp \cap E_j(q_j^*, \theta_j^*)|\ge2$, we have $|\Pp \cap \partial E_j(q_j', \theta_j)| \ge 2$.
\end{definicao}

Next, we introduce a notation that helps us characterize angles which given an ellipse rotated by it and two points, it is possible to find a center for the ellipse, such that it contains both points.

\begin{definicao}\label{def:feasible_angle}
	Let $E$ be the coverage region of an ellipse and $u, v \in \R^2$. An angle $\theta \in [0, \pi)$ is said to be $(E, u, v)$-feasible if there is $q \in \R^2$ such that $\{u, v\} \subset \partial E(q, \theta)$.
	In addition to that, given an instance  $(\Pp, \Ww, \Rr)$ of MCER, the set of $(E_j, u, v)$-feasible angles is referred to as

	\begin{equation}
	\Phi_j(u, v) := \{\theta \in [0, \pi) : \theta \textnormal{ is a } (E_j,u,v)\textnormal{-feasible angle}\}.
	\end{equation}
	We also define $\tilde{\Phi}_j(u,v)$ as the angle which makes $E_j$'s major-axis be parallel to the line that passes through $u$ and $v$. Note that if $\Phi_j(u,v) \neq \emptyset$, then $\tilde{\Phi}_j(u,v) \in \Phi_j(u,v)$ as the longest segment that crosses an ellipse is its major-axis.
\end{definicao}

In \autoref{fig:feasible-angle} two examples for \autoref{def:feasible_angle} are shown. The example with a solid border shows two given points on two different ellipses rotated by $\pi/4$, making $\pi/4$ a $(E, u, v)$-feasible angle.
The other example, with a dashed border, presents a case where the two points cannot be on the ellipse rotated by $\pi/2$, no matter where it is placed; because of that, $\pi/2$ is said to be a non $(E, u, v)$-feasible angle.

\begin{figure}[H]
	\centering
	\caption{A $(E, u, v)$-feasible angle and a not $(E, u, v)$-feasible angle.}
	\includegraphics[scale=.28]{tex/figures/feasible-angle2}
	\fautor
	\label{fig:feasible-angle}
\end{figure}

Following that, we introduce a lemma that is responsible for connecting the developments of this chapter with the results of \autoref{chapter:e3p}.
This lemma makes it possible to describe a type of solution which, for sure, is part of the equivalence class of any optimal solution.
It states that, for any ellipse that covers more than two points in a given optimal solution, an equivalent solution exists with at least one of the two properties:
\begin{itemize}
	\item The ellipse contains at least three points.
	\item The ellipse contains two points for any feasible angle.
\end{itemize}

\begin{lema}\label{lema:3pnts}
Let $(Q^*, \Theta^*)$ be an optimal solution of an instance $(\Pp, \Ww, \Rr)$ of MCER; \\$j\in\{1, \dots, m\}$, such that $|\Pp \cap E_j(q_j^*, \theta_j^*)|\ge2$; $(Q', \Theta')\in\Pi(Q^*, \Theta^*)$; and $\{u, v\}\subset \partial E_j(q_j', \theta_j')$.\\
If, for all $(\hat{Q}, \hat{\theta})$ equivalent solution of $(Q^*, \Theta^*)$, $|\Pp\cap \partial E_j(\hat{q}_j, \hat{\theta}_j)| < 3$, then for all $\theta\in\Phi_j(u,v)$, there exists $q\in\R^2$, such that $\{u, v\} \subset \partial E_j(q, \theta)$ and $\Pp \cap E_j(q^*_j, \theta^*_j) = \Pp \cap E_j(q, \theta)$.

\end{lema}

\begin{proof}
	According to \autoref{lema:mce_2b}, there exists $\{u, v\} \subset \Pp \cap E_j(q^*_j, \theta^*_j)$, such that an equivalent optimal solution $(Q', \Theta')$ exists with $u$ and $v$ on the border of $E_j(q_j', \theta_j^*)$. Therefore, $\theta_j^*\in\Phi_j(u,v)$.

	Now suppose that $u$ and $v$ have the same $y$-coordinate, that is, the angle between them is $0$. If they do not, a rotation can be applied to make them have the same $y$-coordinate. Then, the first thing we are proving is that $\Phi_j(u, v) = [0, 2\alpha]$ for a specific case that any instance can be transformed into using translation and rotation on every element of $\Pp$.

	In \autoref{chapter:definitions}, a function $L\colon \R \to \R_{\ge0}$ was defined in \autoref{eq:function-l}. This function takes the angular coefficient $m\in\R$ and, considering the family of lines parallel to the one described by $y=mx$, returns the maximum squared distance between two intersection points of a line in that family and an axis-parallel ellipse centered at the origin.

	To use those results here, we need to consider the ellipse to be fixed at the origin and axis-parallel, and consider the problem of rotating and translating the points in $\Pp$ instead.

	Let $\theta \in [0, \pi]\setminus\{\pi/2\}$, and $u', v'$ be the points $u, v$ after a rotation by $\theta$. Then, if $L(\tan{\theta}) \ge ||v-u||_2^2$, it is possible to apply a translation to $u',v'$, such that they end up on the fixed ellipse. This means that it is possible to find an angle of rotation and a center to place $E_j$, such that it has $u, v$ on its border.

	Now we use some properties of function $L$ whose details are given in \autoref{chapter:definitions}.
	Defining $l(\theta)=L(\tan{\theta})$, with $l:[0, \pi]\setminus\{\pi/2\}$, we can say that

	\begin{itemize}
		\item $l$ is decreasing in $[0, \pi/2)$ because $L$ is decreasing in $[0, \infty)$. Therefore, if there is $\alpha\in[0, \pi/2)$, such that $l(\alpha) = ||v-u||_2^2$, then $l(\theta)>||v-u||_2^2$, for $\theta\in(\alpha, \pi/2)$. That implies $[0, \alpha] \subset \Phi_j(u,v)$.
		\item $l(\theta) = l(\pi-\theta)$ because $L$ is an even function. Therefore, if there is $\alpha\in[0, \pi/2)$, such that $l(\alpha) = ||v-u||_2^2$, then $l(\theta)>||v-u||_2^2$, for $\theta\in(\pi/2,\pi-\alpha)$. That implies $[\pi-\alpha, \pi] \subset \Phi_j(u,v)$.
	\end{itemize}
	We then conclude that $\Phi_j(u, v) = [0, \alpha]\cup [\pi-\alpha, \pi]$, and, of course, in the case that there is no $\alpha\in[0, \pi/2)$, such that $l(\alpha)=||v-u||_2^2$, we have $\Phi_j(u,v)=[0, \pi]$.
	From that, if we rotate every point in $\Pp$ by $\pi-\alpha$, we obtain $\Phi_j(u,v)=[0, 2\alpha]$.

	With this result in hand, we can use a continuity argument to complete our proof as follows.
	Let $\delta : \Phi_j(u,v) \mapsto \R^2$ be a continuous function which takes an angle $\theta\in\Phi_j(u,v)$ and returns a center, such that $\{u,v\} \subset \partial E_j(\delta(\theta), \theta)$, and, from solution $(Q', \Theta')$, $\delta(\theta_j') = q_j'$. Notice that, in general, for any angle in $\Phi_j(u,v)$, there are two possible centers that make $\{u,v\} \subset \partial E_j(\delta(\theta), \theta)$ (see \autoref{fig:feasible-angle} for an example), however, imposing $\delta(\theta_j') = q_j'$ makes $\delta$ be a well-defined function. This is shown in \autoref{fig:lema-3-points} where $\delta$ is plotted for the whole interval $\Phi_j(u,v)$.

	Let $w\in \Pp \setminus \{u,v\}$, then we define $f_w  : \R^2 \times [0, \pi) \mapsto \R_{\ge0}$ to be a function that takes a center $q \in \R^2$ and an angle of rotation $\theta\in [0, \pi)$, and returns the elliptical distance between $w$ and $E_j(q, \theta)$ minus $1$ as defined by the left-hand-side of \autoref{eq:rotated_ellipse}. Keep in mind that, for all $w\in \Pp \cap E_j(q_j^*, \theta_j^*)\setminus \{u,v\}$, we have that $f_w(q_j', \theta_j') < 0$ as they are covered by the $j$-th ellipse.

	Then, to evaluate $f_w$ for every center and feasible angle that maintains $u$ and $v$ on $E_j$'s border, we introduce a new function $g_w\colon \Phi_j(u,v) \mapsto \R_{\ge0}$, which is defined as $g_w(\theta) = f_w(\delta(\theta), \theta)$.
	As $f_w$ and $\delta$ are both continuous functions, $g_w$ is also continuous.

	Therefore, for any $\theta\in\Phi_j(u,v)$, if a point $w \in \Pp \cap E_j(q_j^*, \theta_j^*)\setminus\{u,v\}$ is not covered by $E_j(\delta(\theta), \theta)$, it must have $g_w(\theta)>0$. Then, by continuity, another angle $\bar{\theta} \in \Phi_j(u,v)$ must exist, such that $g_w(\bar{\theta})=1$, which means that $w\in \partial E_j(\delta(\hat{\theta}), \hat{\theta})$, contradicting the hypothesis. The same can be said about the case where there exists an angle $\theta \in \Phi_j(u,v)$, such that a point $w \in \Pp \setminus E_j(q_j^*, \theta_j^*)$ enters the coverage of $E_j(\delta(\theta), \theta)$.
\end{proof}

%\begin{figure}[H]%
%	\centering%
%	\caption{Every ellipse that contains two points for a fixed angle of %rotation.}
%	\includegraphics[scale=.36]{tex/figures/2point-ellipse}
%	\fautor
%	\label{fig:2point-ellipse}
%\end{figure}

What \autoref{lema:3pnts} states is that, for every ellipse in an instance of MCER, unless an equivalent optimal solution with three points on it exists, the angle of rotation can practically be ignored. Because of that, it will be shown that we can construct a CLS for each ellipse which is finite and also contains an optimal solution.

In \autoref{fig:lema-3-points}, a visualization of \autoref{lema:3pnts} is presented.
An initial optimal solution is given by the dashed-border ellipse and its center, represented by a star point. From it, the continuous function $\delta$ is defined by moving the ellipse through the rotation angles in $\Phi_j(u,v)$ while maintaining $u, v$ on it. Ten angles were chosen from $\Phi_j(u,v)$ to be shown in \autoref{fig:lema-3-points}, among those were $0$ and $\max\{\Phi_j(u,v)\}$; their corresponding ellipses are displayed with solid-line borders.
Consistently with \autoref{lema:3pnts}, the points in $\Pp \setminus \{u, v, w\}$ stay within the ellipse's cover for any angle of rotation, and, for point $w$, there exists an angle, such that it is on the ellipse.

\begin{figure}[H]
	\centering
	\caption{A visualization of \autoref{lema:3pnts}.}
	\includegraphics[scale=.9]{tex/figures/lema-3-points}
	\fautor
	\label{fig:lema-3-points}
\end{figure}

Let $(Q^*, \Theta^*)$ be any optimal solution of an instance $(\Pp, \Ww, \Rr)$ of MCER. We will define a set of solutions $\Omega(\Pp, \Ww, \Rr)$ and show that there exists an equivalent solution $(Q', \Theta')$ to  $(Q^*, \Theta^*)$, such that $(Q', \Theta') \in \Omega(\Pp, \Ww, \Rr)$. This is the same thing as showing that $\Omega(\Pp, \Ww, \Rr)$ for sure contains an optimal solution for the instance $(\Pp, \Ww, \Rr)$.
Before that, we introduce a definition for the CLS of every ellipse, which is then used to construct the set of solutions $\Omega(\Pp, \Ww, \Rr)$.

\begin{definicao}\label{def:Sj}
	Let $(\Pp, \Ww, \Rr)$ be an instance of MCER. Then, for all $j\in\{1, \dots, m\}$, we define the \sigla{CLS}{Candidate Locations Set} of the $j$-th ellipse as $S_j = S_j^{(1)} \cup S_j^{(2)} \cup S_j^{(3)}$ with
	\begin{align}
		S_j^{(1)} &= \bigcup_{u \in \Pp} \{(u, 0)\}\\
		S_j^{(2)} &= \bigcup_{\{u, v\} \subset \Pp} \{(q, \tilde{\Phi}_j(u,v))\in \R^2\times\R: \{u,v\} \subset \partial E_j(q, \tilde{\Phi}_j(u,v))\}\\
		S_j^{(3)} &= \bigcup_{\{u, v, w\} \subset \Pp} \{(q, \theta)\in \R^2\times\R: \{u, v, w\} \subset \partial E_j(q, \theta)\}.
	\end{align}
\end{definicao}

This definition breaks the construction of the CLS $S_j$ into three separated cases. The first one, $S_j^{(1)}$, represents solutions where the $j$-th ellipse covers only one point. The second one, $S_j^{(2)}$, takes into account solutions where the $j$-th ellipse covers at least two points, and no equivalent solution with three points on the ellipse exists. The last case, $S_j^{(3)}$, considers solutions where there exists an equivalent one with three points on the $j$-th ellipse. Following this, we move forward and introduce the main result of this section.

\begin{theorem}\label{th:mcer}
	Let $(\Pp, \Ww, \Rr)$ be an instance of MCER, and $\Omega(\Pp, \Ww, \Rr)$ be a set of solutions defined as 
	\begin{equation*}\label{eq:omega}
	\Omega(\Pp, \Ww, \Rr) = \{(Q, \Theta)\in \R^{2m}\times\R^m \colon (q_j, \theta_j) \in S_j \textnormal{ for all } j\in\{1, \dots, m\}\},
	\end{equation*}
Then there exists an optimal solution $(Q^*, \Theta^*) \in \Omega(\Pp, \Ww, \Rr)$, and $|\Omega(\Pp, \Ww, \Rr)|=\bigO(n^{3m})$.
\end{theorem}
\begin{proof}
	The first thing to notice is that $\Omega(\Pp, \Ww, \Rr)$ is defined as the combination of every possible solution from each CLS. To prove that it contains an optimal solution $(Q^*, \Theta^*)$, we only need to prove that for all $j\in\{1, \dots, m\}$, there exists $(q_j, \theta_j)\in S_j$, such that $\Pp \cap E_j(q_j^*, \theta_j^*) \subset \Pp \cap E_j(q_j, \theta_j)$. That is, we only need to show that the CLS of every ellipse contains a center and angle of rotation that makes the ellipse cover the same points (possibly some additional ones) that it covers in an optimal solution. To do that, we use \autoref{lema:3pnts} and break the possible optimal solutions into three cases.

	 In the first case, we consider solutions where the $j$-th ellipse covers less than one point, that is, $|\Pp \cap E_j(q_j^*, \theta_j^*)|\le1$. It is possible to see that $S_j^{(1)}$ takes this possibility into account as it includes in $\Omega(\Pp, \Ww, \Rr)$ every solution that has an ellipse centered at a point from $\Pp$. From that, we can also conclude that $|S_j^{(1)}| = n$.

	In the second case, we consider solutions where the $j$-th ellipse covers at least two points, and no equivalent solution exists, such that three points are on it. This case is addressed by \autoref{lema:3pnts}, which says that an optimal solution $(Q^*, \Theta^*)$ of this type has equivalent solutions with two points $u, v \in \Pp \cap E_j(q_j^*, \theta_j^*)$ on the ellipse, for any angle of rotation $\theta_j \in \Phi_j(u, v)$.
	As $\tilde{\Phi}_j(u,v) \in \Phi_j(u,v)$, we have that there exists $(q_j, \theta_j) \in S_j^{(2)}$, such that $\Pp \cap E_j(q_j^*, \theta_j^*) = \Pp \cap E_j(q_j, \theta_j)$. 
	Moreover, in \autoref{section:ellipses_intersection}, we discuss the problem of determining the intersections between two ellipses. This problem is equivalent to the problem of determining every center and angle of rotation that puts two points on an ellipse, which is the problem used in the definition of $S_j^{(2)}$ for every pair of points. In \autoref{section:ellipses_intersection}, it is shown that this problem has at most two solutions. Therefore, we have that $|S_j^{(2)}| = 2\binom{n}{2}$.

	For the last case, we are left with solutions where the $j$-th ellipse covers more than two points, and there exists an equivalent solution with three points on it. 
	As $S_j^{(3)}$ contains every center and angle of rotation that puts three points on the $j$-th ellipse, an equivalent solution for this case is present in the set of solutions $\Omega(\Pp, \Ww, \Rr)$. Also, in \autoref{lema:3pnts} it is stated that the number of centers and angles of rotation that make an ellipse contain three given points is at most six. Therefore, we have that $|S_j^{(3)}| = 6\binom{n}{3}$.

	Finally, as $|S_j| \le |S_j^{(1)}| +  |S_j^{(2)}| +  |S_j^{(3)}| = \bigO(n^{3})$, it follows that $|\Omega(\Pp, \Ww, \Rr)| = |S_1|\times \dots \times |S_m| = \bigO(n^{3m})$.
\end{proof}

\section{An algorithm for MCER}

In this section we describe an algorithm for MCER that does a complete search on the CLS of each ellipse.
Firstly, in \autoref{algoritmo:mcer1}, we present a procedure called CLS-MCER which returns the CLS for an ellipse with shape parameters $(a, b)$. Then, in \autoref{algoritmo:mcer} we describe the procedure that returns an optimal solution for MCER.

Let $E$ be the coverage region of an ellipse with shape parameters $(a, b)$. We assume that in \autoref{algoritmo:mcer1}, the procedure $e2p(u, v, a, b)$ returns every $(q, \tilde{\Phi}_j(u,v)) \in \R^2\times[0, \pi)$, such that, $\{u, v\}\subset \partial E(q, \tilde{\Phi}_j(u,v))$. That is, this procedure returns every location for the ellipse with shape parameters $(a, b)$, such that its angle of rotation is $\tilde{\Phi}_j(u,v)$, and the points $u$, $v$ are on the ellipse.

This procedure can be defined using the results of \autoref{section:ellipses_intersection}.

\begin{algoritmo}
	\caption{An algorithm that constructs a CLS for a given ellipse.}\label{algoritmo:mcer1}
	\begin{algorithmic}[1]
		\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, and an ellipse's shape parameters $(a, b)$.}

		\Ensure{A CLS for the ellipse with shape parameters $(a, b)$ considering the demand set $\Pp$.}

		\item[]

		\Procedure{CLS-MCER}{$\Pp, a, b$}
		\State $S \gets \{\}$

		\For {$u \in \Pp$}
		\State $S \gets S \cup \{(u, 0)\}$
		\EndFor

		\For {$\{u, v\} \in \Pp$}
		\State $S \gets S \cup e2p(u, v, a, b)$
		\EndFor

		\For{$\{u, v, w\} \in \Pp$}
		\State $S \gets S \cup e3p(u, v, a, b)$ \Comment{Defined in \autoref{algoritmo:e3p}.}
		\EndFor

		\State \Return $S$
		\EndProcedure
	\end{algorithmic}
\end{algoritmo}

After that, we define \autoref{algoritmo:mcer} which takes an instance $(\Pp, \Ww, \Rr)$ of MCER, and returns an optimal solution for it.
Even though it is based on \autoref{th:mcer}, the set of solutions $\Omega(\Pp, \Ww, \Rr)$ is not explicitly built in \autoref{algoritmo:mcer}. Instead, a complete search is done by backtracking the CLS of every ellipse returned by procedure CLS-MCER defined in \autoref{algoritmo:mcer1}.

Two procedures are defined in \autoref{algoritmo:mcer}. The first one, called $MCER$, returns an optimal solution for an instance $(\Pp, \Ww, \Rr)$ using the second procedure $MCER_{bt}$. This second procedure is responsible for the backtracking and takes two additional parameters $j\in\{1, \dots, m\}$, which represents the index of the ellipse that $MCER_{bt}$ is currently processing, and $Z\subset \Pp$, which represents the set of points that have not been covered by the ellipses with indexes $1, \dots, j-1$.

\begin{algoritmo}
	\caption{Algorithm for MCER}\label{algoritmo:mcer}
	\begin{algorithmic}[1]
		\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, a list of weights $\Ww=\{w_1, \dots, w_n\}$, and a list of shape parameters $\Rr=\{(a_1, b_1), \dots, (a_m, b_m)\}$.}

		\Ensure{An optimal solution for the given instance of MCER.}

		\item[]
		\Procedure{$MCER$}{$\Pp, \Ww, \Rr$}
		\State \Return $MCER_{bt}(\Pp, \Ww, \Rr, 1)$
		\EndProcedure

		\item[]

		\Procedure{$MCER_{bt}$}{$Z, \Ww, \Rr, j$}
		%\If{$j = m+1$}
		%\State \Return $0$
		%\EndIf

		%\State $ans \gets 0$
		\State $(q_{j}^*, \dots, q_m^*); (\theta_{j}^*, \dots, \theta_m^*) \gets (0, \dots 0); (0, \dots, 0)$ \Comment{Setting to $0$ as a default value.}

		\State $S_j \gets \textnormal{CLS-MCER}(Z, a_j, b_j)$

		\ForAll{$(q_j, \theta_j) \in S_j$}
		%\State $Cov \gets \Pp \cap E_j(q_j, \theta_j)$

		\If{$j < m$}
		\State $(q_{j+1}, \dots, q_m); (\theta_{j+1}, \dots, \theta_m) \gets  MCER_{bt}(Z\setminus Cov, \Ww, \Rr, j+1)$
		\EndIf

		%\State $Cov \gets \bigcup_{k=j}^m \Pp \cap E_k(q_k, \theta_k)$
		\If {$w(\bigcup_{k=j}^m \Pp \cap E_k(q_k, \theta_k)) > w(\bigcup_{k=j}^m \Pp \cap E_k(q_k^*, \theta_k^*))$}
			%\State $ans \gets w(Cov)$
			%\State $((q_{j}^*,\theta_{j}^*); \dots; (q_m^*, \theta_m^*)) \gets ((q_{j},\theta_{j}); \dots; (q_m, \theta_m))$

			\State $(q_{j}^*, \dots, q_m^*); (\theta_{j}^*, \dots, \theta_m^*) \gets (q_{j}, \dots, q_m); (\theta_{j}, \dots, \theta_m)$
		\EndIf
		\EndFor

		\State \Return $(q_{j}^*, \dots, q_m^*); (\theta_{j}^*, \dots, \theta_m^*)$
		\EndProcedure
	\end{algorithmic}
\end{algoritmo}

\begin{corolario}
	\autoref{algoritmo:mcer} takes $\bigO(n^{3m})$ operations and returns an optimal solution for an instance of MCER.
\end{corolario}

\begin{proof}
	For every $j\in\{1, \dots, m\}$, unless $Z=\{\}$, when choosing the center and angle of rotation for the $j$-th ellipse, \autoref{algoritmo:mcer} does not consider any $(q_j, \theta_j) \in \R^2 \times \R$, such that $Z \cap E_j(q_j, \theta_j) = \{\}$. Apart from those solutions, which are non-optimal, \autoref{algoritmo:mcer} considers every solution in $\Omega(\Pp, \Ww, \Rr)$. As evaluating each solution can be done in $\bigO(n)$, we get the overall runtime complexity of $\bigO(n^{3m+1})$ which is $\bigO(n^{3m})$.
\end{proof}

In \autoref{fig:AB120}, a solution returned by \autoref{algoritmo:mcer} for the instance AB120 taken from \citeonline{andreta} is displayed. The exact method developed by \citeonline{andreta}, for this instance, could not obtain an optimal solution within the established time limit, however, comparing with the solution obtained by our algorithm, their heuristic method does find an optimal solution. In the next chapter, we give more details about the solutions found by our algorithm, along with the proposal of some new instances for MCER.

\begin{figure}[!htb]
	\centering
	\caption{An optimal solution for the instance AB120.}
	\includegraphics[scale=.5]{tex/figures/AB120}
	\fautor
	\label{fig:AB120}
\end{figure}

\subsection{Adding facility cost}

In this section, we consider the extended version of MCER where each facility has a cost assigned to it and exactly $k$ of them must be used in a solution. We refer to this version of the problem as \sigla{MCER-$k$}{Maximum Covering by Ellipses with Rotation and a $k$-constraint}.

An instance of MCER-$k$ has the same parameters as MCER plus a list of costs \mbox{$\Cc:=\{c_1, \dots, c_m\}$}, with $c_j\in\R_{>0}$ being the cost of the $j$-th facility; and $k\in\mathbb{N}$, $k\le m$.

A solution for MCER-$k$ is given by $(I, Q, \Theta)$, with $I:=\{i_1, \dots, i_k\}\subset \{1, \dots, m\}$; \\\mbox{$Q:=(q_1, \dots, q_k)\in\R^{2k}$}, with $q_j$ being the center of the $i_j$-th ellipse; and \mbox{$\Theta:=(\theta_1, \dots, \theta_k) \in [0, \pi)^k$}, with $\theta_j$ being the angle of rotation of the $i_j$-th ellipse. An optimal solution of MCER-$k$ is given by the optimization problem

\begin{equation*}
	\max_{I, Q, \Theta} w\left(\bigcup_{j=1}^k \Pp \cap E_{i_j}(q_j, \theta_j)\right).
\end{equation*}

Then, in the same way that it is done in \autoref{chapter:mce}, we introduce \autoref{algoritmo:mcer-k} for MCER-$k$ that uses \autoref{algoritmo:mcer} for every $I:=\{i_1, \dots, i_k\} \subset \{1, \dots, m\}$. Therefore, \autoref{algoritmo:mcer-k} returns an optimal solution for MCER-$k$ in $\bigO(\binom{m}{k}n^{3k}) = \bigO(n^{4m})$ time.

\begin{algoritmo}
	\caption{Algorithm for MCER-$k$}\label{algoritmo:mcer-k}

	\begin{algorithmic}[1]
		\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, a list of weights $\Ww=\{w_1, \dots, w_n\}$, a list of shape parameters $\Rr=\{(a_1, b_1), \dots, (a_m, b_m)\}$, a list of costs $\Cc=\{c_1, \dots, c_m\}$, and $k\in \mathbb{N}$.}
		\Ensure{An optimal solution for MCER-$k$.}

		\item[]

		\Procedure{MCER-$k$}{$\Pp, \Ww, \Rr, \Cc, k$}
		\State $I^* = \{i_1^*, \dots, i_k^*\}\gets \{1, \dots, k\}$
		\State $Q^* = (q_1^*, \dots, q_k^*) \gets (0, \dots, 0)$
		\State $\Theta^* = (\theta_1^*, \dots, \theta_k^*) \gets (0, \dots, 0)$

		\ForAll{$I=\{i_1, \dots, i_k\} \subset \{1, \dots, m\}$}

		\State $\Rr' \gets \{(a_j, b_j) \in \Rr: j \in I\}$
		\State $(q_1, \dots, q_k); (\theta_1, \dots, \theta_k) \gets MCER(\Pp, \Ww, \Rr')$

		\If{$w(\bigcup_{j=1}^k \Pp \cap E_{i_j}(q_j, \theta_j)) - \sum_{j\in I} c_j > w(\bigcup_{j=1}^k \Pp \cap E_{i_j^*}(q_j^*, \theta_j^*)) - \sum_{j\in I^*}c_{j}$}
		\State $Q^* \gets (q_1, \dots, q_k)$
		\State $\Theta^* \gets (\theta_1, \dots, \theta_k)$
		\State $I^* \gets I$
		\EndIf
		\EndFor

		\State \Return $I^*, Q^*, \Theta^*$
		\EndProcedure
	\end{algorithmic}
\end{algoritmo}

\subsection{Improvements and implementation details}

Some improvements can be made in the implementation of \autoref{algoritmo:mcer}.
Instead of calling procedure CLS-MCER every time in the backtracking routine, it is suggested that it is done outside in a pre-processing phase.

Let $(Q, \Theta)$ and $(Q', \Theta')$ be two solutions of MCER. If, for any $j \in \{1, \dots, m\}$, we have $\Pp \cap E_j(q_j', \theta_j') \subset \Pp \cap E(q_j, \theta_j)$, solution $(Q', \Theta')$ can be dismissed. This can be done in the CLS-MCER procedure in \autoref{algoritmo:mcer1}.

In our implementation we use the same approach used in \citeonline{andreta}. When constructing the CLS for the $j$-th ellipse, a tree-like data structure is utilized to verify if a solution $(q, \theta)$ covers a subset of another solution or not.

Keeping an upper-bound for the solution during the backtracking in \autoref{algoritmo:mcer} can also provide a good improvement in practice. Let $z_j = \max_{(q, \theta) \in S_j} w(\Pp \cap E_j(q, \theta))$, and $OPT$ the value of an optimal solution. When the algorithm chooses $(q_j, \theta_j) \in S_j$ as the solution for the $j$-th ellipse, we have

\begin{equation}
OPT \le w\left(\bigcup_{k=1}^{j} \Pp \cap E_k(q_k, \theta_k)\right) + \sum_{k=j+1}^m z_k.
\end{equation}
This upper-bound is easy to compute as $\{z_1, \dots, z_m\}$ can be pre-processed, and the weight of every point covered so far in the backtrack can be obtained by adding another parameter to the $MCER$ procedure.

In the next chapter, we present the results of numerical experiments for MCER and MCER-$k$ and compare them with the results published in \citeonline{andreta}.