### 2020-03-04 defesa/tex/chapter/tables.tex

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|crrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{5}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & \# nodes & \# sol. leaves & CLS-MCE & Total\\
				\hline
				\input{tex/tables/table_mcek_cm.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCE-$k$ for instances CM1-CM9 obtained by \autoref{algoritmo:mce-k}.}
		\label{tab:mce-results-cm}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|ccrrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{6}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\# E3P&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & subproblems & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%
				%%%

				\hline
				\input{tex/tables/table_mcerk_cm.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCER-$k$ for instances CM1-CM9 obtained by \autoref{algoritmo:mcer-k}.}
		\label{tab:mcer-results-cm}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{0.85\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|crrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{5}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & \# nodes & \# sol. leaves & CLS-MCE & Total\\
				\hline
				\input{tex/tables/table_mcek_ab_1.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCE-$k$ for instances AB001-AB060 obtained by \autoref{algoritmo:mce-k}.}
		\label{tab:mce-results-ab1}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{0.85\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|crrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{5}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header
				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & \# nodes & \#sol. leaves & CLS-MCE & Total\\
				\hline
				\input{tex/tables/table_mcek_ab_2.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCE-$k$ for instances AB061-AB120 obtained by \autoref{algoritmo:mce-k}.}
		\label{tab:mce-results-ab2}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|ccrrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{6}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\# E3P&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & subproblems & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%
				%%%

				\hline
				\input{tex/tables/table_mcerk_ab_1.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCER-$k$ for instances AB001-AB060 obtained by \autoref{algoritmo:mcer-k}.}
		\label{tab:mcer-results-ab1}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|ccrrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{6}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\# E3P&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & subproblems & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcerk_ab_2.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCER-$k$ for instances AB061-AB120 obtained by \autoref{algoritmo:mcer-k}.}
		\label{tab:mcer-results-ab2}
	\end{center}
\end{table}

%\section{New instances}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|crrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{5}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcek_ta_1.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCE-$k$ for instances TA001-TA007 obtained by \autoref{algoritmo:mce-k}.}
		\label{tab:mce-results-ta1}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|ccrrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{6}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\# E3P&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & subproblems & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcerk_ta_1.tex}
				\hline

			\end{tabular}
		}
		\caption{Solutions of MCER-$k$ for instances TA001-TA007 obtained by \autoref{algoritmo:mcer-k}.}
		\label{tab:mcer-results-ta1}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|crrrr|}
	\hline
	\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{5}{c|}{Performance metrics}\\
	\hline

	%%% Second line of header

	\multirow{2}{*}{Name} & 
	\multirow{2}{*}{$n$} & 
	\multirow{2}{*}{$m$} & 
	\multirow{2}{*}{$k$} & 
	Selected & 
	\multirow{2}{*}{Income} & 
	CLS size&
	\multicolumn{2}{c}{Backtracking Tree} & 
	\multicolumn{2}{c|}{\centering CPU Time (s)}\\
	& & & & \centering Ellipses & & $|S_k|$ & \# nodes & \#sol leaves & CLS-MCER & Total\\

	%%%%
	%%%%

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcek_ta_2.tex}
				\hline
				%\input{tex/tables/table_mcerk_ta_3.tex}
				%\hline

			\end{tabular}
		}
		\caption{Solutions of MCE-$k$ for instances TA008-TA022 obtained by \autoref{algoritmo:mce-k}.}

		\label{tab:mce-results-ta2}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|ccrrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{6}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\# E3P&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & subproblems & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcerk_ta_2.tex}
				\hline
				%\input{tex/tables/table_mcerk_ta_3.tex}
				%\hline

			\end{tabular}
		}
		\caption{Solutions of MCER-$k$ for instances TA008-TA022 obtained by \autoref{algoritmo:mcer-k}.}

		\label{tab:mcer-results-ta2}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|crrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{5}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcek_ta_4.tex}
				\hline
				%\input{tex/tables/table_mcerk_ta_3.tex}
				%\hline

			\end{tabular}
		}
		\caption{Solutions of MCE-$k$ for instances TA008-TA022 obtained by \autoref{algoritmo:mce-k}.}

		\label{tab:mce-results-ta3}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|ccrrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{6}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\# E3P&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & subproblems & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcerk_ta_4.tex}
				\hline
				%\input{tex/tables/table_mcerk_ta_3.tex}
				%\hline

			\end{tabular}
		}
		\caption{Solutions of MCER-$k$ for instances TA008-TA022 obtained by \autoref{algoritmo:mcer-k}.}

		\label{tab:mcer-results-ta3}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|crrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{5}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcek_ta_5.tex}
				\hline
				%\input{tex/tables/table_mcerk_ta_3.tex}
				%\hline

			\end{tabular}
		}
		\caption{Solutions of MCE-$k$ for instances TA008-TA022 obtained by \autoref{algoritmo:mce-k}.}

		\label{tab:mce-results-ta4}
	\end{center}
\end{table}

\begin{table}[!htb]
	\begin{center}
		\resizebox{\textwidth}{!}{%

			\begin{tabular}{|cccc|cr|ccrrrr|}
				\hline
				\multicolumn{4}{|c|}{Instance} & \multicolumn{2}{c|}{Optimal Solution} & \multicolumn{6}{c|}{Performance metrics}\\
				\hline

				%%% Second line of header

				\multirow{2}{*}{Name} & 
				\multirow{2}{*}{$n$} & 
				\multirow{2}{*}{$m$} & 
				\multirow{2}{*}{$k$} & 
				Selected & 
				\multirow{2}{*}{Income} & 
				CLS size&
				\# E3P&
				\multicolumn{2}{c}{Backtracking Tree} & 
				\multicolumn{2}{c|}{\centering CPU Time (s)}\\
				& & & & \centering Ellipses & & $|S_k|$ & subproblems & \# nodes & \#sol leaves & CLS-MCER & Total\\

				%%%%
				%%%%

				\hline
				\input{tex/tables/table_mcerk_ta_5.tex}
				\hline
				%\input{tex/tables/table_mcerk_ta_3.tex}
				%\hline

			\end{tabular}
		}
		\caption{Solutions of MCER-$k$ for instances TA008-TA022 obtained by \autoref{algoritmo:mcer-k}.}

		\label{tab:mcer-results-ta4}
	\end{center}
\end{table}