### 2020-03-11 defesa/article/e3p.tex
wip

In this section, we introduce the problem of determining every location, here defined as the center and angle of rotation, of an ellipse with fixed shape parameters, such that it contains three given points.
This problem comes up in the development of an algorithm for MCER in the next section. 
It is important to point out that no prior studies were found on it, or even on related problems.
We propose an algorithm for it that involves determining the eigenvalues of a $6\times6$ complex matrix. We also analyze its efficiency in terms of numerical accuracy and display some solutions that it was able to obtain.

\subsection{Problem definition}

Given the shape parameters of an ellipse $(a, b)$, $a > b >0$, and three points $u, v, w \in \R^2$, let $E\colon \R^2\times [0, \pi)\to\R^2$ be the coverage region of an ellipse with shape parameters $(a, b)$, we refer to the problem of obtaining $(q, \theta) \in \R^2\times [0, \pi)$, such that $\{u, v, w\} \subset \partial E(q, \theta)$ as Ellipse by Three Points Problem (E3P).
Because of its application here in our work, we are only interested in a method that can obtain every solution of E3P.

\subsection{Transforming E3P into a circle problem}

Initially, E3P is a problem with three unknown variables: $(q_x, q_y)$, and $\theta$. However, we can reduce this number to only one, as it will be shown, it is possible to obtain $q$ from $\theta$.
Let us assume that point $u$ is at the origin. If it is not, a simple translation by $-u$ applied to the three points can be made to put $u$ at the origin.
Assume as well that $(q, \theta)$ is a solution of E3P.

Applying a rotation of $-\theta$ to the coordinate system makes the ellipse in the original solution become axis-parallel.
Then, that ellipse can be transformed into a circle of radius $b$ by squeezing the $x$-axis by $b/a$. This two-step transformation can be written as a function $\varphi\colon \R^2 \to \R^2$ defined as

\begin{equation}\label{eq:trpnts}
\varphi(p, \theta)=\left[\begin{array}{cc}
\frac{b}{a}&0\\
0&1
\end{array}\right]
\left[\begin{array}{cc}
\cos{\theta}&\sin{\theta}\\
-\sin{\theta}&\cos{\theta}
\end{array}\right]\left[\begin{array}{c}
p_x\\
p_y
\end{array}\right].
\end{equation}

An example of this transformation can be seen in \autoref{fig:circumscribed-circle}. As $\varphi^{-1}$ is well-defined, instead of solving E3P, we can work with the univariate problem of determining an angle of rotation $\theta \in [0, \pi)$ that makes the triangle with vertices $\varphi(u, \theta), \varphi(v, \theta), \varphi(w, \theta)$ be circumscribed in a circle of radius $b$. To make the notation less cluttered, we denote by $\Lambda(\theta)$ the triangle with vertices $\varphi(u, \theta), \varphi(v, \theta), \varphi(w, \theta)$.

Let $A(\theta)$ denote the area of the triangle $\Lambda(\theta)$. Using the formula given in \cite[p.~189]{johnson1960} for the radius of the circumscribed circle of a triangle, we can obtain a function $\xi \colon [0, \pi) \to \R$ given by
\begin{equation}\label{eq:circumscribed_circle_b}
\xi(\theta) = 16b^2A(\theta)^2 - \norm{\varphi(v, \theta)}^2\norm{\varphi(w, \theta)}^2\norm{\varphi(v, \theta)-\varphi(w, \theta)}^2,
\end{equation}
whose roots determine solutions of E3P.

\begin{figure}
	\centering

	\includegraphics[scale=.3]{../tex/figures/circumscribed-circle}
	\caption{Transforming a solution of E3P into a solution of the circumradius problem.}
	\label{fig:circumscribed-circle}
\end{figure}

\begin{lem}\label{lema:e3p}
	E3P has at most six solutions.
\end{lem}

\begin{pf}
	The first thing to notice is that $\xi$ is a real trigonometric polynomial of degree $6$. 
	Its term of highest degree is the multiplication of the norms $\norm{\varphi(v, \theta)}^2\norm{\varphi(w, \theta)}^2\norm{\varphi(v, \theta)-\varphi(w, \theta)}^2$. In \cite[p.~150]{powell}, where a definition of real trigonometric polynomial is also given, it is stated that a $n$-degree real trigonometric polynomial can have up to $2n$ roots in $[0, 2\pi)$. Therefore, E3P has at most $12$ solutions in $[0, 2\pi]$.
	Half of these solutions, though, are duplicated as ellipses are symmetric to their axis.
\end{pf}

\subsection{Converting $\xi$ into a polynomial}

In \cite[p.~195]{horn}, a theorem is presented stating that for every univariate polynomial of degree $n$, there exists a companion matrix, which is a $n\times n$ matrix, such that its eigenvalues are the zeros of that polynomial. Finding every eigenvalue of a matrix can be done using the QR algorithm, which runs in $\bigO(n^3)$ and uses $\bigO(n^2)$ memory (a very complete introduction to it can be found in \cite{watkins:2008}). For example, for a degree-$4$ polynomial $\sum_{k=0}^4 a_k x^4$, a possible companion matrix is given by

 \begin{equation*}
\left[\begin{array}{ccccc}
0 & 1 & 0 & 0\\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1\\
-\dfrac{a_0}{a_4} & -\dfrac{a_1}{a_4} & -\dfrac{a_2}{a_4} & -\dfrac{a_3}{a_4}
\end{array}\right].
\end{equation*}
In practice, we can use the very well-known LAPACK software library to obtain the eigenvalues of a matrix; for more information about LAPACK see \cite{lapack}.
This approach works for both real or complex polynomials, because of that, based on \cite{weidner}, we describe a way of converting $\xi$ into a complex polynomial.

By using the two identities $\cos{\theta} = (e^{i\theta} + e^{-i\theta})/2$, and $\sin{\theta} = (e^{i\theta} - e^{-i\theta})/(2i)$, which relate trigonometric functions with complex numbers in the unit circle $\mathbb{S}=\{z\in \mathbb{C}\colon |z|=1\}$, we can rewrite $\xi$ as a function of the new variable $z=e^{i\theta}\in\mathbb{S}$. In \cite{weidner}, it is stated that this substitution when utilized for the task of determining the roots of a real trigonometric polynomial does not yield loss of accuracy.

As $\xi$ is a real trigonometric polynomial of degree $6$, $z$ appears with exponents from $-6$ up to $6$. Multiplying $\xi$ by $z^6$ and extending the domain of $z$ to $\mathbb{C}$, we obtain a complex polynomial $g(z)=\sum_{k=0}^{12} c_k z^k$, for some $c_0, \dots, c_{12} \in \mathbb{C}$. In practice, symbolic computation can be used to obtain the actual coefficients of $g$ in terms of an instance of E3P.

Let $angle \colon \mathbb{C} \colon [0, 2\pi)$ be a function that returns the angle of a complex number, then given a root $\hat{z}$ of $g$, if $|\hat{z}|=1$ and $angle(\hat{z}) \in [0, \pi)$, then $\hat{\theta} = angle(\hat{z})$ is a root of $\xi$. Observing that for any $z\in \mathbb{C}$, $angle(-z)=\pi+angle(z)$, which is the same symmetry ellipses have, we conclude that $g(-z)=g(z)$, implying that all the odd-degree coefficients of $g$ are zero. Using the substitution $y=z^2$, we can obtain a degree-$6$ polynomial $f(y)=\sum_{k=1}^6 c_{2k} y^k$. Then, from a root $\hat{y}$ of $f$, we obtain a root $\hat{\theta} = angle(\hat{y})/2$ of $\xi$.
Lastly, let $C(\theta)\in \R^2$ be the center of the circumcircle of $\Lambda(\theta)$, then from a root $\hat{\theta}$ of $\xi$, we obtain the solution $(\varphi^{-1}(C(\hat{\theta})) + u, \hat{\theta})$ of E3P.

It worth mentioning that a pattern on the coefficients of $f$ was identified, and maybe it can be used for further improvements. Analyzing the polynomials produced for several instances, we observed that $c_k = \overline{c_{6-k}}$, for all $k=0\dots 6$.

\subsection{Choosing a precision constant}

We developed an experiment to choose a precision constant for comparing if a root of $f$ is in the unit circle. The implementation was coded in C++, and LAPACK's ZGEEV was utilized to obtain the eigenvalues of the companion matrix of the polynomial $f$, more information about the implementation is given in \autoref{section:implementation}. For the experiment, we defined $K\in\R$, $K>0$, and consider instances with ellipse's shape parameters $(K, \frac{K}{2})$, for $K\in\{10^j\colon j=0,\dots,10\}$.

The experiment considers instances of E3P where the three points are the vertices of an ellipse rotated by $\theta\in[0, \pi)$. Such instances only have one solution, and therefore, roots with multiplicity greater than one are expected, which can be seen as a special case.
For each value of $K$, we ran the algorithm for $100$ instances generated randomly by sampling $\theta$ according to a uniform distribution. For each instance, we took the root $\hat{z}$ which produced the closest solution to the priorly known one. Then, for each $K$, as it can be seen in \autoref{fig:e3p_known_sols2}, we considered the maximum and the average distance to the unit circle $|1-|\hat{z}||$; and, as it can be seen in \autoref{fig:e3p_known_sols1}, the maximum and average error $|f(\hat{z})|$.

From this experiment, we decided to adopt a precision constant of $10^{-6}$ to consider a root of $f$ to be in the unit circle, and as an additional check, we adopted a precision constant of $10^{-9}$ to consider a root to be a solution of E3P.

\begin{figure}

	\begin{subfigure}{.5\textwidth}
		\centering

		\includegraphics[scale=.5]{../tex/figures/e3p_known_sols2}
		\caption{}
		\label{fig:e3p_known_sols2}
	\end{subfigure}
	\begin{subfigure}{.5\textwidth}
	\centering

	\includegraphics[scale=.5]{../tex/figures/e3p_known_sols1}
	\caption{}
	\label{fig:e3p_known_sols1}
	\end{subfigure}
\caption{(a) shows the maximum and average distance to the unit circle $|1-|\hat{z}||$. (b) shows the maximum and average error $|f(\hat{z})|$, where $\hat{z}$ is a root of $f$ returned by our algorithm.}
\end{figure}

\subsection{Instances with four and six solutions}

Any instance of E3P, as stated by \autoref{lema:e3p} can have up to six solutions. At first, though, this bound seemed to be loose as for randomly generated instances, we were not able to obtain instances with more than two solutions.

After some investigation, though, we were able to construct some four-solution instances (an example is displayed in \autoref{fig:e3p_4sols}). An interesting property of those solutions is that every one of them has their three points form an isosceles triangle.

\begin{figure}
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[scale=.28]{../tex/figures/e3p_4sols}
		\caption{An instance of E3P with four solutions.}
		\label{fig:e3p_4sols}
	\end{subfigure}
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[scale=.28]{../tex/figures/e3p_6sols}
		\caption{An instance of E3P with six solutions.}
		\label{fig:e3p_6sols}
	\end{subfigure}
\end{figure}

Obtaining six-solution instances, on the other hand, was done by taking a particular case of the isosceles-triangle approach.
As it can be seen in \autoref{fig:e3p_6sols}, the three points on every one of the six ellipses' border form an equilateral triangle.

It should be pointed out that neither non-isosceles instances with four solutions nor non-equilateral instances with six solutions could be found. Further investigating these possible properties of E3P is left as future work.