### 2020-03-12 defesa/article/mcer.tex

The version of PMCLP where the facilities are ellipses that can be freely rotated was first introduced in \cite{andreta} where an exact and a heuristic method were developed for it. In comparison with MCE, this problem introduces a new variable that is responsible for determining the rotation angle of every ellipse, making MCER a more challenging problem. We propose an algorithm for MCER which is able to obtain optimal solutions for every instance proposed in \cite{andreta} including the ones its exact method could not, and its heuristic obtained non-optimal ones.