### 2020-03-16 defesa/article/implementation.tex
wip

All the algorithms were implemented using the C++ language, compiled with g++ (G++ 6.3.0) with the optimization flag -O3. The actual code is available in \url{https://sites.icmc.usp.br/andretta/tedeschi-2020/}.