### 2020-03-21 defesa/article/conclusion.tex

In this work, we studied two problems of maximum planar covering using ellipses, with most of the substantial results being for the less previously-studied problem where the ellipses can be freely rotated, which we referred to as MCER. 
We devoted most of our work for the development of an exact algorithm for this problem, which depended on the development of an algorithm for a never-studied-before geometric subproblem, which we referred to as E3P. The algorithm proposed for this last problem involves determining the eigenvalues of a six by six complex matrix to obtain the roots of a degree-six polynomial. We believe that further properties can be discovered which would allow the development of a simpler method to obtain the roots of the polynomial whose roots are solutions of E3P.

An exact algorithm was also developed for the other problem where the ellipses had to be axis-parallel. We took a different approach compared to the methods proposed by \cite{canbolat, andreta}, and based our algorithm entirely on the model proposed for the Euclidean PMCLP in \cite{church:1984}.

After giving improvement suggestions for the implementation of both algorithms, we analyzed several numerical experiments that were made using instances from \cite{canbolat, andreta}, and some new ones. Both algorithms performed really well, obtaining optimal solutions for every previously known instance, and only timing out for the really challenging ones from the set of newly constructed ones.