### 2020-03-25 defesa/article/conclusion.tex
wip

In this work, we developed new algorithms for two planar maximum covering by ellipses problems.
After giving improvements suggestions, and implementation details, we analyzed numerical experiments showing the efficiency of both algorithms for the instances of previous works, and for new ones with larger demand and facility sets.

We devoted most of our work to the development of an exact algorithm for MCER.
In the midst of that, a never-studied-before subproblem came up, for which we also proposed an algorithm, which involved determining the roots of a complex polynomial of degree six.
We believe that further exploring this new problem could be the subject of future work as we observed that its solutions, except for one case, always come in pairs; and that this property seems to be directly connected with the complex polynomial, as its roots seem to come in conjugate pairs. These properties, if proved to be true, might be used in the development of a more efficient algorithm.