### 2020-07-19 paper/article.tex
finsh some corrections

\documentclass[3p, 11pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% The graphicx package provides the includegraphics command.

%\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{subcaption}
\usepackage{siunitx}
\sisetup{group-separator = {,}}

\usepackage{multirow}
\usepackage{cancel}

\journal{EJOR}

%\usepackage{xcolor}
\usepackage[dvipsnames]{xcolor}

\usepackage[mathscr]{eucal}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Co}{\mathbb{Co}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\Pp}{\mathscr{P}}
\newcommand{\Cc}{\mathscr{C}}
\newcommand{\E}{\mathscr{E}}
\newcommand{\F}{\mathscr{F}}
\newcommand{\Ww}{\mathscr{W}}
\newcommand{\Rr}{\mathscr{R}}
\newcommand{\norm}[2][2]{\left\lVert#2\right\rVert_{#1}}
\newcommand{\bigO}{\mathscr{O}}
\newtheorem{thm}{Theorem}
\newtheorem{prp}{Proposition}
\newtheorem{lem}{Lemma}
\newdefinition{rmk}{Remark}
\newdefinition{definition}{Definition}
%\newproof{pf}{Proof}
%\renewcommand\qedsymbol{$\blacksquare$}

\bibliographystyle{plainnat}

\RequirePackage{algorithm, algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\algorithmautorefname}{Algorithm}
\newcommand{\definitionautorefname}{Definition}
\newcommand{\thmautorefname}{Theorem}
\newcommand{\prpautorefname}{Proposition}
\newcommand{\lemmaautorefname}{Lemma}
\newcommand{\lemautorefname}{Lemma}
\newcommand{\sigla}[2]{#2 (#1)}
\newcommand{\citeonline}[1]{\cite{#1}}

\begin{document}

	\begin{frontmatter}

		%% Title, authors and addresses

		\title{New Exact Algorithms for Planar Maximum Covering Location by Ellipses Problems\tnoteref{t1}}
		\tnotetext[t1]{This paper is the results of the research
			project funded by CAPES and FAPESP.}

		\author[add1]{Danilo Tedeschi\corref{cor1}}
		\ead{danilo.tedeschi@usp.br}

		\author[add1]{Marina Andretta}
		\ead{andretta@icmc.usp.br}

		\address[add1]{Department of Applied Mathematics and Statistics, Institute of Mathematical and Computer Sciences, University of São Paulo, Avenida Trabalhador São-carlense, 400, Centro, 13566-590, São Carlos, SP, Brazil.}

		\cortext[cor1]{Corresponding author.}
\end{frontmatter}
\clearpage
%\begin{frontmatter}
%		\begin{abstract}

\section*{Abstract}
			%% Text of abstract

			{\color{Red}
		Planar Maximum Covering Location by Ellipses is an optimization problem where one wants to choose the location of ellipses given their major and minor axis to cover demand points, maximizing a function depending on the value of covered points.	
		}We propose new exact algorithms for two versions of this problem, one where the ellipses have to be parallel to the coordinate axis, and another where they can be freely rotated. 
			Besides finding optimal solutions for previously published instances, including the ones where no optimal solution was known, both algorithms proposed by us were able to obtain optimal solutions for some new larger instances 	with up to seven hundred demand points and five ellipses.
\subsection*{Keywords}
			Combinatorial optimization, planar maximal covering location problem, planar covering by ellipses, exact algorithms.
			%% keywords here, in the form: keyword \sep keyword

			%% MSC codes here, in the form: \MSC code \sep code
			%% or \MSC[2008] code \sep code (2000 is the default)

%		\end{keyword}

%	\end{frontmatter}

\section{Introduction}\label{section:introduction}

\input{introduction.tex}

\section{Definition}\label{section:definition}
\input{definition.tex}

\section{An Algorithm for MCE}\label{section:mce}

\input{mce.tex}

\section{Determining Every Center and Angle of Rotation of An Ellipse Given Its Shape and Three Points that It Must Contain}\label{section:e3p}

\input{e3p.tex}

\section{An Algorithm for MCER}\label{section:mcer}

\input{mcer.tex}

\section{Improvements}\label{section:improvements}
\input{improvements.tex}
%\section{Reducing the CLS size}\label{section:reducing_cls}
%\input{reducing_cls.tex}

%\section{Prunning the Backtracking Tree}\label{section:prunning}
%\input{prunning.tex}

\section{Implementation Details}\label{section:implementation}
\input{implementation.tex}

\section{Numerical Experiments}\label{section:numerical}
\input{numerical.tex}

\section{Conclusion}\label{section:conclusion}
\input{conclusion.tex}

\clearpage
%\section*{References}
	%% New version of the num-names style
	%\bibliographystyle{natdin}
	\bibliography{../defesa/references}

\end{document}