### 2020-08-17 paper/acknowledgments.tex
wip

%We are very grateful for the editorial process conducted by this journal, every comment, and suggestion sent to us by the reviewers were very insightful and detailed; without them, we could not have improved our work the way we did, reaching the quality it is now.

We would like to thank the reviewers for their useful suggestions, that greatly improved this work.
We would also like to express our gratitude to the funding institutions Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES), and Fundação de Amparo à Pesquisa do Estado de São Paulo (FAPESP) (processes 2013/07375-0 and 2016/01860-1) that made this research possible.